---
title: "Land cover ratio: example on RTP area"
output: html_document
date: "2023-11-09"
---

Load function

```{r, message=F}
pkgs <- c(
  "terra", "tidyterra", "exactextractr", "sf",
  "dplyr", "fastDummies", "viridis", "ggplot2",
  "fastDummies"
)
sapply(pkgs, library, character.only = TRUE)
source("../R/create_prediction_grid.R")
```

Open NC shapefile

```{r}
nc <- vect(paste0(
  "../input/NC_county_boundary/",
  "North_Carolina_State_and_County_Boundary_Polygons.shp"
))
same.crs(crs(nc), "EPSG:6543")
plot(nc)
```

## Create prediction grid

Open imperviousness data

```{r}
imp <- rast("../input/NC_imperviousness_2019.tif")
same.crs(crs(imp), "EPSG:4326")
same.crs(crs(imp), "EPSG:5070")
nc_proj <- terra::project(nc, crs(imp))
imp <- terra::mask(imp, nc_proj)
plot(imp)
# check that linear units is in meters:
linearUnits(imp)
```

Create 300m raster with aggregated imperviousness

```{r}
imp_agg <- aggregate(imp, fact = 10, fun = "mean") # aggregation at 300m
imp_agg <- terra::mask(imp_agg, nc_proj)
plot(imp_agg > 20)
```

```{r}
# 300m prediction grid as raster
pred_rast <- imp_agg

# 300m prediction grid as vector
pred_vect <- as.points(pred_rast)
```

```{r}
pred_rast
extend(pred_rast, c(3, 3))
```

## Add land cover on RTP area example

Open land cover data and create NLCD dataframe with all information

```{r}
lc <- rast("../input/NC_nlcd_crs-wgs84.tif")
nc_proj <- terra::project(nc, crs(lc))
lc <- terra::crop(lc, nc_proj)
plot(lc)

nlcd_classes <- list(
  value = c(
    11, 21, 22, 23, 24, 31, 41, 42, 43, 52,
    71, 81, 82, 90, 95
  ),
  class = c(
    "WTR", "OSD", "LID", "MID", "HID",
    "BRN", "DFO", "EFO", "MFO", "SHB",
    "GRS", "PAS", "CRP", "WDW", "EHW"
  ),
  names = c(
    "Open Water",
    "Developed, Open Space",
    "Developed, Low Intensity",
    "Developed, Medium Intensity",
    "Developed, High Intensity",
    "Barren Land",
    "Deciduous Forest",
    "Evergreen Forest",
    "Mixed Forest",
    "Shrub/Scrub",
    "Herbaceous",
    "Hay/Pasture",
    "Cultivated Crops",
    "Woody Wetlands",
    "Emergent Herbaceous Wetlands"
  ),
  col = c(
    "#476ba1", "#decaca", "#d99482", "#ee0000",
    "#ab0000", "#b3aea3", "#68ab63", "#1c6330",
    "#b5ca8f", "#ccba7d", "#e3e3c2", "#dcd93d",
    "#ab7028", "#bad9eb", "#70a3ba"
  )
)
nlcd_classes <- as.data.frame(nlcd_classes)
levels(lc) <- nlcd_classes
```

#### Select RTP area

```{r}
lat <- c(35.6, 36.11, 36.11, 35.6)
lon <- c(-79.19, -79.19, -78.39, -78.39)
ext_rtp <- vect(cbind(lon, lat), type = "polygons", crs = "EPSG:4326")
ext_rtp <- terra::project(ext_rtp, crs(pred_rast))
pred_rast_rtp <- terra::crop(pred_rast, ext_rtp)
plot(pred_rast_rtp)
pred_vect_rtp <- terra::as.points(pred_rast_rtp)
```

#### Function to compute NLCD class ratio per buffer

```{r}
# x is a SpatVector with crs unit in meter
# nlcd is a 30m-resolution raster
compute_ndld_class_ratio0 <- function(spvect, nlcd, width = 150) {
  # create circle buffers with 150m radius
  bufs_pol <- terra::buffer(spvect, width = width)

  # crop nlcd raster
  extent <- terra::ext(bufs_pol)
  nlcd_cropped <- terra::crop(nlcd, extent)

  # turn nlcd variable to dummies
  nlcd_cropped_df <- as.data.frame(nlcd_cropped, xy = TRUE)
  nlcd_cropped_df <- dummy_cols(nlcd_cropped_df)
  nlcd_cropped <- as_spatraster(nlcd_cropped_df,
    xycols = c(1, 2),
    crs = crs(nlcd_cropped)
  )

  # ratio of each nlcd class per buffer
  nlcd_at_bufs <- terra::extract(nlcd_cropped,
    bufs_pol,
    fun = function(x) sum(x, na.rm = TRUE) / length(x)
  )
  new_spvect <- cbind(spvect, nlcd_at_bufs)
  new_spvect <- new_spvect[names(new_spvect)[grepl(
    "class_",
    names(new_spvect)
  )]]
  names(new_spvect) <- sub("class_", "", names(new_spvect))
  return(new_spvect)
}
```

Application to RTP area

```{r}
start_time <- Sys.time()
pred_nlcd_ratio_vect <- compute_ndld_class_ratio0(pred_vect_rtp, lc)
end_time <- Sys.time()
end_time - start_time
pred_nlcd_ratio_df <- as.data.frame(pred_nlcd_ratio_vect, geom = "XY")
pred_nlcd_ratio_df <- cbind(
  pred_nlcd_ratio_df[, c("x", "y")],
  pred_nlcd_ratio_df[
    ,
    !(colnames(pred_nlcd_ratio_df) %in% c("x", "y"))
  ]
)
pred_nlcd_ratio_rast <- as_spatraster(pred_nlcd_ratio_df, xycols = c(1, 2))

nlcd_classes

ggplot() +
  geom_spatraster(data = pred_nlcd_ratio_rast) +
  facet_wrap(~lyr, ncol = 4) +
  scale_fill_viridis()
```

#### Final optimized function

It uses exactextractr library and is much more efficient.

```{r}
source("../R/process_spatial_covariates.R")
```

```{r}
start_time <- Sys.time()
pred_nlcd_ratio_vect <- compute_ndld_class_ratio(pred_vect_rtp, lc)
end_time <- Sys.time()
end_time - start_time

pred_nlcd_ratio_df <- as.data.frame(pred_nlcd_ratio_vect, geom = "XY")
pred_nlcd_ratio_df <- cbind(
  pred_nlcd_ratio_df[, c("x", "y")],
  pred_nlcd_ratio_df[
    ,
    !(colnames(pred_nlcd_ratio_df) %in% c("x", "y"))
  ]
)
pred_nlcd_ratio_rast <- as_spatraster(pred_nlcd_ratio_df,
  xycols = c(1, 2)
)

nlcd_classes

ggplot() +
  geom_spatraster(data = pred_nlcd_ratio_rast) +
  facet_wrap(~lyr, ncol = 4) +
  scale_fill_viridis()
```
