---
title: "map-nc-land-covariates"
output: html_document
date: "2023-07-18"
---

Libraries

```{r, results='hide', echo=FALSE}
options("sp_evolution_status" = 2) # use sf instead of rgdal and rgeos in sp
library(sp)
library(terra)
library(ncdf4)

# -- for timeseries
library(lubridate)
library(zoo)
library(xts)

# --
library(ggplot2)
library(ggspatial)
library(tidyterra)
library(maditr)
library(tidyverse)
library(viridis)
```

```{r}
input_path <- "../input/"
```

#### NC shapefile

```{r, warning=FALSE}
nc_poly <- paste0("NC_county_boundary/",
                  "North_Carolina_State_and_County_Boundary_Polygons.shp")
nc_borders <- vect(paste0(input_path, nc_poly))
```

```{r, echo=FALSE}
crs(nc.borders)
```

## NOAA aws analysis

```{r}
aws_file <- "NC-AWS-NOAA-dailysummary-20220601-20220831.csv"
aws <- data.table::fread(paste0(input_path, aws_file))
length(unique(aws$STATION))
aws <- aws[which(!(is.na(aws$TMAX) | is.na(aws$TMIN))), ]
aws$DATE <- as.Date(aws$DATE, format = "%Y-%m-%d")
length(unique(aws$STATION))
```

Daily max temperatures

```{r, message=FALSE}
aws_ts_tx <- maditr::dcast(aws[, c("DATE", "STATION", "TMAX")],
                           DATE ~ STATION) %>%
  as.xts()
plot(aws_ts_tx)

nb_na_tx <- lapply(aws_ts_tx, FUN = function(x) sum(is.na(x)))
nb_na_tx <- as.data.frame(do.call(rbind, nb_na_tx))
nb_na_tx <- cbind(STATION = rownames(nb_na_tx), nb_na_tx)
rownames(nb_na_tx) <- seq_len(nrow(nb_na_tx))
names(nb_na_tx)[names(nb_na_tx) == "V1"] <- "nb_na_tx"
```

Daily min temperatures

```{r, message=FALSE}
aws_ts_tn <- maditr::dcast(aws[, c("DATE", "STATION", "TMIN")],
                           DATE ~ STATION) %>%
  as.xts()
boxplot(t(aws_ts_tx))

nb_na_tn <- lapply(aws_ts_tn, FUN = function(x) sum(is.na(x)))
nb_na_tn <- as.data.frame(do.call(rbind, nb_na_tn))
nb_na_tn <- cbind(STATION = rownames(nb_na_tn), nb_na_tn)
rownames(nb_na_tn) <- seq_len(nrow(nb_na_tn))
names(nb_na_tn)[names(nb_na_tn) == "V1"] <- "nb_na_tn"
```

Daily mean temperatures

```{r, message=FALSE}
aws_ts_tm <- maditr::dcast(aws[, c("DATE", "STATION", "TAVG")],
                           DATE ~ STATION) %>%
  as.xts()
plot(aws_ts_tm)

nb_na_tm <- lapply(aws_ts_tm, FUN = function(x) sum(is.na(x)))
nb_na_tm <- as.data.frame(do.call(rbind, nb_na_tm))
nb_na_tm <- cbind(STATION = rownames(nb_na_tm), nb_na_tm)
rownames(nb_na_tm) <- seq_len(nrow(nb_na_tm))
names(nb_na_tm)[names(nb_na_tm) == "V1"] <- "nb_na_tm"
```

Map NOAA aws with less than 5% missing data

```{r , message=FALSE}
stations <- unique(aws[, c("STATION", "NAME", "LATITUDE",
                           "LONGITUDE", "ELEVATION")])
stations <- list(stations, nb_na_tx, nb_na_tn, nb_na_tm) %>%
  reduce(full_join, by = "STATION")
aws <- list(aws, nb_na_tx, nb_na_tn, nb_na_tm) %>%
  reduce(full_join, by = "STATION")

# -- turn into a sf object to reproject CRS
stations <- st_as_sf(stations, coords = c("LONGITUDE", "LATITUDE"))
st_crs(stations) <- "EPSG:4326"
stations <- st_transform(stations, crs(nc.borders))
stations <- data.frame(stations)

ggplot() +
  geom_sf(
    data = st_as_sf(nc.borders), aes(geometry = geometry),
    colour = "grey", fill = NA
  ) +
  geom_sf(
    data = stations[which(stations$nb_na_tx <= 5), ],
    aes(geometry = geometry),
    size = .5, fill = NA
  ) +
  scale_color_discrete(name = "")
```

Map minimum temperatures on 2022-07-07 (one of the hottest day)

```{r, message=FALSE}
# -- turn into a sf object to reproject CRS
aws <- st_as_sf(aws, coords = c("LONGITUDE", "LATITUDE"))
st_crs(aws) <- "EPSG:4326"
aws <- st_transform(aws, crs(nc.borders))
aws <- data.frame(aws)

ggplot() +
  geom_sf(
    data = st_as_sf(nc.borders), aes(geometry = geometry),
    colour = "grey", fill = NA
  ) +
  geom_sf(
    data = aws[which(aws$DATE == as.Date("2022-07-07")), ],
    aes(geometry = geometry, color = (TMIN - 32) * 5 / 9),
    size = 2
  ) +
  scale_color_viridis(option = "H")
```

## Spatial covariates maps

Create entent for RTP area if we want to plot a zoom

```{r}
lat <- c(35.6, 36.11, 36.11, 35.6)
lon <- c(-79.19, -79.10, -78.39, -78.39)
ext_rtp <- vect(cbind(lon, lat), type = "points", crs = "EPSG:4326")
ext_rtp <- project(ext_rtp, crs(nc.borders))
ext(ext_rtp)
```

### Digital Elevation Model (DEM)

Open all .TIF files and merge them to create a single DEM file for the entire NC.

```{r, eval=FALSE}
dir <- "../input/NC_DEM/"
files <- list.files(path = dir, full.names = TRUE)
rasters <- lapply(files, FUN = rast)
rasters <- sprc(rasters)
dem <- merge(rasters)
terra::writeRaster(dem, "../input/NC-DEM.tif")
```

Plot DEM

```{r, eval=FALSE}
dem <- rast("../input/NC-DEM.tif")
dem <- project(dem, crs.meters)
dem_rtp <- crop(dem, ext_rtp)

p_dem <- ggplot() +
  geom_spatraster(data = dem) +
  geom_sf(
    data = st_as_sf(nc.borders), aes(geometry = geometry),
    colour = "grey", linewidth = .3, fill = NA
  ) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "m"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE),
    na.value = NA
  ) +
  labs(
    fill = "",
    title = "Digital Elevation Model (DEM)"
  ) +
  annotation_scale(
    location = "bl", pad_x = unit(1, "cm"),
    pad_y = unit(1, "cm"),
    height = unit(0.30, "cm"),
    text_cex = 1
  ) +
  annotation_north_arrow(
    location = "br",
    which_north = "true",
    pad_x = unit(0.2, "cm"),
    pad_y = unit(0.2, "cm")
  ) +
  theme(
    axis.text = element_text(size = 12, family = "serif"),
    plot.caption = element_text(size = 10, family = "serif"),
    legend.text = element_text(size = 12, family = "serif"),
    legend.title = element_text(size = 12, family = "serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey")
  )

p_dem_rtp <- ggplot() +
  geom_spatraster(data = dem_rtp) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "m"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE),
    na.value = NA
  ) +
  labs(
    fill = "",
    title = "Digital Elevation Model (DEM) Duhram-Raleigh-Chapel Hill"
  ) +
  annotation_scale(
    location = "bl", pad_x = unit(1, "cm"),
    pad_y = unit(1, "cm"),
    height = unit(0.30, "cm"),
    text_cex = 1
  ) +
  annotation_north_arrow(
    location = "br",
    which_north = "true",
    pad_x = unit(0.2, "cm"),
    pad_y = unit(0.2, "cm")
  ) +
  theme(
    axis.text = element_text(size = 12, family = "serif"),
    plot.caption = element_text(size = 10, family = "serif"),
    legend.text = element_text(size = 12, family = "serif"),
    legend.title = element_text(size = 12, family = "serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey")
  )
```

### Imperviousness

```{r}
imp <- rast("../input/NC_imperviousness_2019.tif")
imp <- project(imp, crs.meters)

imp_plot <- ifel(imp == 0, NA, imp)
imp_plot_rtp <- crop(imp_plot, ext_rtp)

p_imp <- ggplot() +
  geom_spatraster(data = imp_plot) +
  geom_sf(
    data = st_as_sf(nc.borders), aes(geometry = geometry),
    colour = "grey", linewidth = .3, fill = NA
  ) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "%"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE),
    na.value = NA
  ) +
  labs(
    fill = "",
    title = "Imperviousness"
  ) +
  annotation_scale(
    location = "bl", pad_x = unit(1, "cm"),
    pad_y = unit(1, "cm"),
    height = unit(0.30, "cm"),
    text_cex = 1
  ) +
  annotation_north_arrow(
    location = "br",
    which_north = "true",
    pad_x = unit(0.2, "cm"),
    pad_y = unit(0.2, "cm")
  ) +
  theme(
    axis.text = element_text(size = 12, family = "serif"),
    plot.caption = element_text(size = 10, family = "serif"),
    legend.text = element_text(size = 12, family = "serif"),
    legend.title = element_text(size = 12, family = "serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey")
  )

p_imp_rtp <- ggplot() +
  geom_spatraster(data = imp_plot_rtp) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "%"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE),
    na.value = NA
  ) +
  labs(
    fill = "",
    title = "Imperviousness Duhram-Raleigh-Chapel Hill"
  ) +
  annotation_scale(
    location = "bl", pad_x = unit(1, "cm"),
    pad_y = unit(1, "cm"),
    height = unit(0.30, "cm"),
    text_cex = 1
  ) +
  annotation_north_arrow(
    location = "br",
    which_north = "true",
    pad_x = unit(0.2, "cm"),
    pad_y = unit(0.2, "cm")
  ) +
  theme(
    axis.text = element_text(size = 12, family = "serif"),
    plot.caption = element_text(size = 10, family = "serif"),
    legend.text = element_text(size = 12, family = "serif"),
    legend.title = element_text(size = 12, family = "serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey")
  )
```

### Tree canopy cover

```{r}
tcc <- rast("../input/NC_tree-canopy-cover_2021.tif")
tcc <- project(tcc, crs.meters)

tcc_plot <- ifel(tcc != 0, tcc, NA)
tcc_plot_rtp <- crop(tcc_plot, ext_rtp)

p_tcc <- ggplot() +
  geom_spatraster(data = tcc_plot) +
  geom_sf(
    data = st_as_sf(nc.borders), aes(geometry = geometry),
    colour = "white", linewidth = .3, fill = NA
  ) +
  scale_fill_whitebox_c(
    palette = "gn_yl",
    direction = -1,
    labels = scales::label_number(suffix = "%"),
    n.breaks = 12,
    limits = c(0, 100),
    guide = guide_legend(reverse = TRUE),
    na.value = NA
  ) +
  labs(
    fill = "",
    title = "Tree canopy cover"
  ) +
  annotation_scale(
    location = "bl", pad_x = unit(1, "cm"),
    pad_y = unit(1, "cm"),
    height = unit(0.30, "cm"),
    text_cex = 1
  ) +
  annotation_north_arrow(
    location = "br",
    which_north = "true",
    pad_x = unit(0.2, "cm"),
    pad_y = unit(0.2, "cm")
  ) +
  theme(
    axis.text = element_text(size = 12, family = "serif"),
    plot.caption = element_text(size = 10, family = "serif"),
    legend.text = element_text(size = 12, family = "serif"),
    legend.title = element_text(size = 12, family = "serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey")
  )

p_tcc_rtp <- ggplot() +
  geom_spatraster(data = tcc_plot_rtp) +
  scale_fill_whitebox_c(
    palette = "gn_yl",
    direction = -1,
    labels = scales::label_number(suffix = "%"),
    n.breaks = 12,
    limits = c(0, 100),
    guide = guide_legend(reverse = TRUE),
    na.value = NA
  ) +
  labs(
    fill = "",
    title = "Tree canopy cover Duhram-Raleigh-Chapel Hill"
  ) +
  annotation_scale(
    location = "bl", pad_x = unit(1, "cm"),
    pad_y = unit(1, "cm"),
    height = unit(0.30, "cm"),
    text_cex = 1
  ) +
  annotation_north_arrow(
    location = "br",
    which_north = "true",
    pad_x = unit(0.2, "cm"),
    pad_y = unit(0.2, "cm")
  ) +
  theme(
    axis.text = element_text(size = 12, family = "serif"),
    plot.caption = element_text(size = 10, family = "serif"),
    legend.text = element_text(size = 12, family = "serif"),
    legend.title = element_text(size = 12, family = "serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey")
  )
```

### Building footprint

Important: it sounds like there are missing tiles in the data. An explanation 
can be found in the paper Heris, M.P., Foks, N., Bagstad, K., and Troy, A., 
2020, A national dataset of rasterized building footprints for the U.S.: U.S. Geological Survey data release, <https://doi.org/10.5066/P9J2Y1WG.>:

"*We also identified systematic gaps in the Microsoft data for some geographic 
areas. These larger gaps seem to have a tile pattern, where aerial photos may 
have been unavailable to the Microsoft building detection algorithm"*

Their computational algorithm is applied to Microsoft released a U.S.-wide 
vector building dataset provided in 2018 but this dataset has missing tiles.

```{r}
build_fp <- rast("../input/NC_building-footprints/NorthCarolina_sum.tif")
build_fp <- project(build_fp, crs.meters)

build_fp_plot <- ifel(build_fp != 0, build_fp, NA)
build_fp_rtp <- crop(build_fp_plot, ext_rtp)

p_build_fp <- ggplot() +
  geom_spatraster(data = build_fp_plot) +
  geom_sf(
    data = st_as_sf(nc.borders), aes(geometry = geometry),
    colour = "white", linewidth = .3, fill = NA
  ) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = ""),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Building footprint (m^2)"
  ) +
  annotation_scale(
    location = "bl", pad_x = unit(1, "cm"),
    pad_y = unit(1, "cm"),
    height = unit(0.30, "cm"),
    text_cex = 1
  ) +
  annotation_north_arrow(
    location = "br",
    which_north = "true",
    pad_x = unit(0.2, "cm"),
    pad_y = unit(0.2, "cm")
  ) +
  theme(
    axis.text = element_text(size = 12, family = "serif"),
    plot.caption = element_text(size = 10, family = "serif"),
    legend.text = element_text(size = 12, family = "serif"),
    legend.title = element_text(size = 12, family = "serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey")
  )

p_build_fp_rtp <- ggplot() +
  geom_spatraster(data = build_fp_rtp) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = ""),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Building footprint (m^2) Duhram-Raleigh-Chapel Hill"
  ) +
  annotation_scale(
    location = "bl", pad_x = unit(1, "cm"),
    pad_y = unit(1, "cm"),
    height = unit(0.30, "cm"),
    text_cex = 1
  ) +
  annotation_north_arrow(
    location = "br",
    which_north = "true",
    pad_x = unit(0.2, "cm"),
    pad_y = unit(0.2, "cm")
  ) +
  theme(
    axis.text = element_text(size = 12, family = "serif"),
    plot.caption = element_text(size = 10, family = "serif"),
    legend.text = element_text(size = 12, family = "serif"),
    legend.title = element_text(size = 12, family = "serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey")
  )
```

### Building height

```{r}
build_h_file <- paste0("NC_building-height-by-block/",
                       "NC_building-heights-by-block.shp")
build_h <- vect(input_path, build_h_file)
build_h <- project(build_h, crs.meters)
build_h_rtp <- crop(build_h, ext_rtp)

p_build_h <- ggplot() +
  geom_spatvector(data = build_h, aes(fill = Height_cat), color = NA) +
  geom_sf(
    data = st_as_sf(nc.borders), aes(geometry = geometry),
    colour = "white", linewidth = .3, fill = NA
  ) +
  scale_fill_manual(
    breaks = c("Low", "Low-medium", "Medium", "Medium-High",
               "High", "Very high"),
    values = c("#38a700", "#d0ff73", "#feebbf", "#ff7f7e",
               "#e60100", "#720000")
  ) +
  labs(
    fill = "",
    title = "Building height by block"
  ) +
  annotation_scale(
    location = "bl", pad_x = unit(1, "cm"),
    pad_y = unit(1, "cm"),
    height = unit(0.30, "cm"),
    text_cex = 1
  ) +
  annotation_north_arrow(
    location = "br",
    which_north = "true",
    pad_x = unit(0.2, "cm"),
    pad_y = unit(0.2, "cm")
  ) +
  theme(
    axis.text = element_text(size = 12, family = "serif"),
    plot.caption = element_text(size = 10, family = "serif"),
    legend.text = element_text(size = 12, family = "serif"),
    legend.title = element_text(size = 12, family = "serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey")
  )

p_build_h_rtp <- ggplot() +
  geom_spatvector(data = build_h_rtp, aes(fill = Height_cat), color = NA) +
  scale_fill_manual(
    breaks = c("Low", "Low-medium", "Medium", "Medium-High",
               "High", "Very high"),
    values = c("#38a700", "#d0ff73", "#feebbf", "#ff7f7e",
               "#e60100", "#720000")
  ) +
  labs(
    fill = "",
    title = "Building height by block Duhram-Raleigh-Chapel Hill"
  ) +
  annotation_scale(
    location = "bl", pad_x = unit(1, "cm"),
    pad_y = unit(1, "cm"),
    height = unit(0.30, "cm"),
    text_cex = 1
  ) +
  annotation_north_arrow(
    location = "br",
    which_north = "true",
    pad_x = unit(0.2, "cm"),
    pad_y = unit(0.2, "cm")
  ) +
  theme(
    axis.text = element_text(size = 12, family = "serif"),
    plot.caption = element_text(size = 10, family = "serif"),
    legend.text = element_text(size = 12, family = "serif"),
    legend.title = element_text(size = 12, family = "serif"),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey")
  )
```

### Maps

```{r}
p_dem
p_tcc
p_imp
p_build_fp
p_build_h

p_dem_rtp
p_tcc_rtp
p_imp_rtp
p_build_fp_rtp
p_build_h_rtp
```

## Meteorological covariates

See era5-reanalysis-processing.Rmd

#### Session info

```{r}
sessionInfo()
```
