---
title: "map-nc-land-covariates"
output: html_document
date: "2023-07-18"
---

Libraries

```{r, results='hide', echo=F}
# -- for spatial data
#library(raster)   # -- old
#library(rgdal)    # -- deprecated in 10/2023
#library(rgeos)    # -- deprecated in 10/2023
#library(maptools) # -- deprecated in 10/2023
options("sp_evolution_status" = 2) # use sf instead of rgdal and rgeos in sp
library(sp)
library(terra)
library(ncdf4)

# -- for timeseries
library(lubridate)
library(zoo)
library(xts)

# -- 
library(ggplot2)
library(ggspatial)
library(tidyterra)
library(maditr)
library(tidyverse)
library(viridis)
```

#### NC shapefile

```{r, warning=FALSE}
nc.borders <- vect("../input/NC_county_boundary/North_Carolina_State_and_County_Boundary_Polygons.shp")
```

```{r, echo=F}
crs(nc.borders)
```

## NOAA aws analysis

```{r}
aws <- data.table::fread("../input/NC-AWS-NOAA-dailysummary-20220601-20220831.csv") 
length(unique(aws$STATION))
aws <- aws[which(!(is.na(aws$TMAX)|is.na(aws$TMIN))),] 
aws$DATE <- as.Date(aws$DATE, format = "%Y-%m-%d")
length(unique(aws$STATION))
```

Daily max temperatures

```{r, message=FALSE}
aws.ts.tx <- maditr::dcast(aws[, c('DATE', 'STATION', 'TMAX')], DATE ~ STATION) %>%
  as.xts() 
plot(aws.ts.tx)

nb.na.tx <- lapply(aws.ts.tx, FUN=function(x) sum(is.na(x)))  
nb.na.tx <- as.data.frame(do.call(rbind, nb.na.tx))
nb.na.tx <- cbind(STATION = rownames(nb.na.tx), nb.na.tx)
rownames(nb.na.tx) <- 1:nrow(nb.na.tx)
names(nb.na.tx)[names(nb.na.tx) == 'V1'] <- 'nb.na.tx'

```

Daily min temperatures

```{r, message=FALSE}
aws.ts.tn <- maditr::dcast(aws[, c('DATE', 'STATION', 'TMIN')], DATE ~ STATION) %>%
  as.xts() 
#plot(aws.ts.tn)
boxplot(t(aws.ts.tx))

nb.na.tn <- lapply(aws.ts.tn, FUN=function(x) sum(is.na(x)))  
nb.na.tn <- as.data.frame(do.call(rbind, nb.na.tn))
nb.na.tn <- cbind(STATION = rownames(nb.na.tn), nb.na.tn)
rownames(nb.na.tn) <- 1:nrow(nb.na.tn)
names(nb.na.tn)[names(nb.na.tn) == 'V1'] <- 'nb.na.tn'
```

Daily mean temperatures

```{r, message=FALSE}
aws.ts.tm <- maditr::dcast(aws[, c('DATE', 'STATION', 'TAVG')], DATE ~ STATION) %>%
  as.xts() 
plot(aws.ts.tm)

nb.na.tm <- lapply(aws.ts.tm, FUN=function(x) sum(is.na(x)))  
nb.na.tm <- as.data.frame(do.call(rbind, nb.na.tm))
nb.na.tm <- cbind(STATION = rownames(nb.na.tm), nb.na.tm)
rownames(nb.na.tm) <- 1:nrow(nb.na.tm)
names(nb.na.tm)[names(nb.na.tm) == 'V1'] <- 'nb.na.tm'
```

Map NOAA aws with less than 5% missing data

```{r , message=FALSE}
stations <- unique(aws[,c('STATION', 'NAME', 'LATITUDE', 'LONGITUDE', 'ELEVATION')])
stations <- list(stations, nb.na.tx, nb.na.tn, nb.na.tm) %>% reduce(full_join, by='STATION')
aws <- list(aws, nb.na.tx, nb.na.tn, nb.na.tm) %>% reduce(full_join, by='STATION')

# -- turn into a sf object to reproject CRS
stations <- st_as_sf(stations, coords=c('LONGITUDE', 'LATITUDE'))
st_crs(stations) <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
stations <- st_transform(stations, crs(nc.borders))
stations <- data.frame(stations)

ggplot() + 
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry),
               colour = "grey", fill = NA) +
  geom_sf(data = stations[which(stations$nb.na.tx<=5),], aes(geometry=geometry),
             size=.5, fill = NA)+
  scale_color_discrete(name="")
```

Map minimum temperatures on 2022-07-07 (one of the hottest day)

```{r, message=F}
# -- turn into a sf object to reproject CRS
aws <- st_as_sf(aws, coords=c('LONGITUDE', 'LATITUDE'))
st_crs(aws) <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
aws <- st_transform(aws, crs(nc.borders))
aws <- data.frame(aws)

ggplot() + 
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "grey", fill = NA) +
  geom_sf(data = aws[which(aws$DATE==as.Date("2022-07-07")),], 
             aes(geometry=geometry, color=(TMIN-32)*5/9), 
             size=2)+
  scale_color_viridis(option = "H")
```

## Spatial covariates maps

Create entent for RTP area if we want to plot a zoom

```{r}
lat <- c(35.6, 36.11, 36.11, 35.6)
lon <- c(-79.19, -79.10, -78.39, -78.39)
ext.rtp <- vect(cbind(lon, lat), type="points", crs="+proj=longlat +datum=WGS84")
ext.rtp <- project(ext.rtp, crs(nc.borders))
ext(ext.rtp)
```

### Digital Elevation Model (DEM)

Open all .TIF files and merge them to create a single DEM file for the entire NC.

```{r, eval=F}
dir <- "../input/NC_DEM/"
files <- list.files(path=dir, full.names = TRUE)
rasters <- lapply(files, FUN = rast)
rasters <- sprc(rasters)
dem <- merge(rasters)
terra::writeRaster(dem, "../input/NC-DEM.tif")
```

Plot DEM

```{r, eval=F}
dem <- rast("../input/NC-DEM.tif")
dem <- project(dem, crs.meters)
dem.rtp <- crop(dem, ext.rtp)

p.dem <- ggplot() +
  geom_spatraster(data = dem) +
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "grey", linewidth=.3, fill = NA) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "m"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE),
    na.value=NA
  ) +
  labs(
    fill = "",
    title = "Digital Elevation Model (DEM)"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)

p.dem.rtp <- ggplot() +
  geom_spatraster(data = dem.rtp) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "m"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE),
    na.value=NA
  ) +
  labs(
    fill = "",
    title = "Digital Elevation Model (DEM) Duhram-Raleigh-Chapel Hill"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
```

### Imperviousness

```{r}
imp <- rast("../input/NC_imperviousness_2019.tif")
imp <- project(imp, crs.meters)

imp.plot <- ifel(imp==0, NA, imp)
imp.plot.rtp <- crop(imp.plot, ext.rtp)

p.imp <- ggplot() +
  geom_spatraster(data = imp.plot) +
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "grey", linewidth=.3, fill = NA) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "%"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE),
    na.value=NA
  ) +
  labs(
    fill = "",
    title = "Imperviousness"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)

p.imp.rtp <- ggplot() +
  geom_spatraster(data = imp.plot.rtp) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "%"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE),
    na.value=NA
  ) +
  labs(
    fill = "",
    title = "Imperviousness Duhram-Raleigh-Chapel Hill"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)

```

### Tree canopy cover

```{r}
tcc <-  rast("../input/NC_tree-canopy-cover_2021.tif")
tcc <- project(tcc, crs.meters)

tcc.plot <- ifel(tcc!=0, tcc, NA)
tcc.plot.rtp <- crop(tcc.plot, ext.rtp)

p.tcc <- ggplot() +
  geom_spatraster(data = tcc.plot) +
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "white", linewidth=.3, fill = NA) +
  scale_fill_whitebox_c(
    palette = "gn_yl",
    direction=-1,
    labels = scales::label_number(suffix = "%"),
    n.breaks = 12,
    limits=c(0,100),
    guide = guide_legend(reverse = T),
    na.value=NA
  ) +
  labs(
    fill = "",
    title = "Tree canopy cover"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)

p.tcc.rtp <- ggplot() +
  geom_spatraster(data = tcc.plot.rtp) +
  scale_fill_whitebox_c(
    palette = "gn_yl",
    direction=-1,
    labels = scales::label_number(suffix = "%"),
    n.breaks = 12,
    limits=c(0,100),
    guide = guide_legend(reverse = T),
    na.value=NA
  ) +
  labs(
    fill = "",
    title = "Tree canopy cover Duhram-Raleigh-Chapel Hill"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
```

### Building footprint

Important: it sounds like there are missing tiles in the data. An explanation can be found in the paper Heris, M.P., Foks, N., Bagstad, K., and Troy, A., 2020, A national dataset of rasterized building footprints for the U.S.: U.S. Geological Survey data release, <https://doi.org/10.5066/P9J2Y1WG.>:

"*We also identified systematic gaps in the Microsoft data for some geographic areas. These larger gaps seem to have a tile pattern, where aerial photos may have been unavailable to the Microsoft building detection algorithm"*

Their computational algorithm is applied to Microsoft released a U.S.-wide vector building dataset provided in 2018 but this dataset has missing tiles.

```{r}
build.fp <-  rast("../input/NC_building-footprints/NorthCarolina_sum.tif")
build.fp <- project(build.fp, crs.meters)

build.fp.plot <- ifel(build.fp!=0, build.fp, NA)
build.fp.rtp <- crop(build.fp.plot, ext.rtp)

p.build.fp <- ggplot() +
  geom_spatraster(data = build.fp.plot) +
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "white", linewidth=.3, fill = NA) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = ""),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Building footprint (m^2)"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)

p.build.fp.rtp <- ggplot() +
  geom_spatraster(data = build.fp.rtp) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = ""),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Building footprint (m^2) Duhram-Raleigh-Chapel Hill"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
```

### Building height

```{r}
build.h <-  vect("../input/NC_building-height-by-block/NC_building-heights-by-block.shp")
build.h <- project(build.h, crs.meters)
build.h.rtp <- crop(build.h, ext.rtp)

p.build.h <- ggplot() +
  geom_spatvector(data = build.h, aes(fill=Height_cat), color=NA) +
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "white", linewidth=.3, fill = NA) +
  scale_fill_manual(
    breaks = c("Low", "Low-medium", "Medium", "Medium-High", "High", "Very high"),
    values = c("#38a700" ,"#d0ff73","#feebbf","#ff7f7e","#e60100", "#720000")
  ) +
  labs(
    fill = "",
    title = "Building height by block"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)

p.build.h.rtp <- ggplot() +
  geom_spatvector(data = build.h.rtp, aes(fill=Height_cat), color=NA) +
  scale_fill_manual(
    breaks = c("Low", "Low-medium", "Medium", "Medium-High", "High", "Very high"),
    values = c("#38a700" ,"#d0ff73","#feebbf","#ff7f7e","#e60100", "#720000")
  ) +
  labs(
    fill = "",
    title = "Building height by block Duhram-Raleigh-Chapel Hill"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)

```

### Maps

```{r}
p.dem
p.tcc
p.imp
p.build.fp
p.build.h

p.dem.rtp
p.tcc.rtp
p.imp.rtp
p.build.fp.rtp
p.build.h.rtp
```

## Meteorological covariates

See era5-reanalysis-processing.Rmd

#### Session info

```{r}
sessionInfo()
```
