---
title: "spatial machine learning"
output: html_document
date: "2023-09-11"
---

Libraries

```{r, message=F}
library(gstat)
library(terra)
library(sf)
library(sftime)
library(tidyverse)
library(ggplot2)
library(ggspatial)
library(gridExtra)
library(tidyterra)
library(rgeos)
library(data.table) # -- for large flat datasets
library(DT)

# -- for timeseries
library(lubridate)
library(xts)

# -- ml libraries
library(caret)
library(glmnet)
library(randomForest)

# -- code style
library(styler)

# -- spatial k-fold CV
library(spatialsample)

# -- tidy tools for quantifying how well model fits to a data set
library(yardstick)
```

My own functions to manipulate data:

```{r}
source("../R/manipulate_spacetime_data.R")
source("../R/open_data_heatmodel.R")
source("../R/ml_check_input_data.R")
source("../R/ml_create_spatial_train_test_sets.R")
source("../R/ml_create_spatial_cross_validation.R")
source("../R/ml_define_models.R")
source("../R/ml_compute_plot_scores_outputs.R")
```

#### Open data on a specific period

```{r}
obs <- fread("../input/NC-monitors-dailysummary-20220601-20220831-space-time-covariates.csv")
nc_borders <- vect("../input/NC_county_boundary/North_Carolina_State_and_County_Boundary_Polygons.shp")

p <- seq(as.Date("2022-07-20"), as.Date("2022-07-22"), by = "1 day")
pred_p <- open_pred_period(p)
obs_p <- obs[date %in% p, ]
```

Define origin and destination projections

```{r}
crs_wgs84 <- "epsg:4326"
crs_nad83 <- "epsg:5070"
```

Project data.tables lon and lat variables

```{r}
pred_p <- project_dt(pred_p, crs_wgs84, crs_nad83)
obs_p <- project_dt(obs_p, crs_wgs84, crs_nad83)
```

Add county to datatables

```{r}
pred_p <- add_nc_county(pred_p, crs_nad83)
obs_p <- add_nc_county(obs_p, crs_nad83)
```

Remove NA in obs_p_sft

```{r}
obs_p <- obs_p[!is.na(obs_p$tmin) & !is.na(obs_p$tmax), ]
```

#### Check data content

```{r}
meta_obs <- c("lon", "lat", "date", "station", "county")
meta_pred <- c("lon", "lat", "date")
predictors <- c("tcc", "imp", "build.fp", "build.h", "dem", "TN12am")
# -- quantitative predictors
predictors_q <- c("tcc", "imp", "build.fp", "dem", "TN12am")
predicted <- "tmin"
check_obs(obs_p, meta_obs, predictors, predicted)
check_pred_grid(pred_p, meta_pred, predictors)
```


#### Prepare train and test samples

Create several types of train and test sets:

-   rndst: select randomly the observations in space and time

-   rnds: select randomly the stations (and keep the whole timeserie)

-   t: seperate train and test by dates (here: remove the last date of the dataset to evaluate capacity of prediction)

-   [area_name]: seperate train and test by selecting stations (and their entire timeserie) according to a contiguous area (group of counties). Several options are possible (test can be urban - mountain - piedmont - sea - plain).

-   net: select monitors per network. Reliable networks: ECONET, CRN, WBAN. Less reliable: COOP and RAWS.

```{r}
set.seed(111)

# -- regular 1-fold
sets_rndst <- create_sets_rndst(obs_p)
train_rndst <- sets_rndst$train
test_rndst <- sets_rndst$test

# -- randomly 1-fold locations out
sets_rnds <- create_sets_rnds(obs_p)
train_rnds <- sets_rnds$train
test_rnds <- sets_rnds$test

# -- randomly 1-fold dates out
test_dates <- as.Date("2022-07-22")
sets_t <- create_sets_t(obs_p, test_dates)
train_t <- sets_t$train
test_t <- sets_t$test

# -- blocks 1-fold locations out

urb <- c("Durham", "Orange", "Chatham", "Wake")
sets_s <- create_sets_s(obs_p, urb)
train_urb <- sets_s$train
test_urb <- sets_s$test

mou <- c("Wilkes", "Ashe", "Watauga", "Caldwell")
sets_s <- create_sets_s(obs_p, mou)
train_mou <- sets_s$train
test_mou <- sets_s$test

pie <- c("Rowan", "Davidson", "Davie", "Forsyth")
sets_s <- create_sets_s(obs_p, pie)
train_pie <- sets_s$train
test_pie <- sets_s$test

pla <- c("Wilson", "Edgecombe", "Nash", "Halifax")
sets_s <- create_sets_s(obs_p, pla)
train_pla <- sets_s$train
test_pla <- sets_s$test

sea <- c("Tyrrell", "Washington", "Dare", "Hyde")
sets_s <- create_sets_s(obs_p, sea)
train_sea <- sets_s$train
test_sea <- sets_s$test

# -- selection with network origin
net <- c("ECONET", "CRN", "WBAN")
sets_net <- create_sets_net(obs_p, net)
train_net <- sets_net$train
test_net <- sets_net$test
```

Plot training et testing sample tile plots and maps

```{r}
df_test <- test_rndst
df_test$set <- "test"
df_train <- train_rndst
df_train$set <- "train"
df <- rbind(df_train, df_test)

ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile() +
  labs(title = "Test on random observations") +
  scale_fill_manual(values = c(train = "lightblue", 
                               test = "tomato4")) +
  my_theme_paper() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
```

```{r}
df_test <- test_rnds
df_test$set <- "test"
df_train <- train_rnds
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile() +
  labs(title = "Test on random monitors") +
  scale_fill_manual(values = c(train = "lightblue", 
                               test = "tomato4")) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
```

```{r}
df_test <- test_t
df_test$set <- "test"
df_train <- train_t
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile() +
  labs(title = "Test on last day") +
  scale_fill_manual(values = c(train = "lightblue", 
                               test = "tomato4")) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
```

```{r}
df_test <- test_urb
df_test$set <- "test"
df_train <- train_urb
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile() +
  scale_fill_manual(values = c("train" = "lightblue", test = "tomato4")) +
  labs(title = "Test on urban county") +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
ggplot(df, aes(x = lon, y = lat, color = set)) +
  geom_point() +
  scale_color_manual(values = c("train" = "lightblue", test = "tomato4")) +
  labs(title = "Test on urban counties") +
  coord_equal()
```

```{r}
df_test <- test_mou
df_test$set <- "test"
df_train <- train_mou
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile() +
  scale_fill_manual(values = c("train" = "lightblue", test = "tomato4")) +
  labs(title = "Test on mountain counties") +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
ggplot(df, aes(x = lon, y = lat, color = set)) +
  geom_point() +
  scale_color_manual(values = c("train" = "lightblue", test = "tomato4")) +
  labs(title = "Test on mountain counties") +
  coord_equal()
```

```{r}
df_test <- test_pie
df_test$set <- "test"
df_train <- train_pie
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile()+
  scale_fill_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on Piedmont counties")+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()
    )  
ggplot(df, aes(x = lon, y = lat, color = set)) +
  geom_point()+
  scale_color_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on Piedmont counties")+
  coord_equal()
```

```{r}
df_test <- test_pla
df_test$set <- "test"
df_train <- train_pla
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile()+
  scale_fill_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on plain counties")+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()
    )  
ggplot(df, aes(x = lon, y = lat, color = set)) +
  geom_point()+
  scale_color_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on plain counties")+
  coord_equal()
```

```{r}
df_test <- test_sea
df_test$set <- "test"
df_train <- train_sea
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile()+
  scale_fill_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on sea counties")+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()
    )  
ggplot(df, aes(x = lon, y = lat, color = set)) +
  geom_point()+
  scale_color_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on sea counties")+
  coord_equal()
```

#### Prepare spatial k-fold CV

```{r}
obs_p_sf <- create_sf(obs_p, "epsg:5070")
sp_samples <- create_sp_fold(obs_p_sf)
```

Plots 

```{r}
cluster_llo <- spatial_clustering_cv(obs_p_sf, v = 10)
block_llo <- spatial_block_cv(obs_p_sf, v = 10)
net_llo <- spatial_leave_location_out_cv(obs_p_sf,
  group = network,
  v = 6
)
rnd_llo <- spatial_leave_location_out_cv(obs_p_sf,
  group = station,
  v = 10
)
autoplot(cluster_llo)
autoplot(block_llo)
autoplot(net_llo)
autoplot(rnd_llo)
```

#### Fit cross validation 

Formula definition

```{r}
(formula <- as.formula(paste(predicted, " ~ ", 
                             paste(predictors, collapse= "+"))))
(formula_q <- as.formula(paste(predicted, " ~ ", 
                             paste(predictors_q,
                                   collapse= "+"))))
```

```{r}
cv_fit <- compute_preds_lm_cv(sp_samples)
cv_rmse <- compute_rmse_cv(cv_fit, predicted=tmin)
cv_rmse
```

Plot RMSE per fold

```{r}
map_rmse_cv(cv_fit, cv_rmse)
```


```{r}
map_res_cv(cv_fit, tmin)
```
```{r}
plot_reg_cv(cv_fit, tmin)
```


## Linear regression

Check gaussian distribution

```{r}
ggplot(obs_p) +
  geom_histogram(aes(tmin), binwidth = 1) +
  xlim(9, 30)

ggplot(obs_p) +
  geom_histogram(aes(tcc), binwidth = 10)

ggplot(obs_p) +
  geom_histogram(aes(log(tcc)))

ggplot(obs_p) +
  geom_histogram(aes(dem), binwidth = 100)

ggplot(obs_p) +
  geom_histogram(aes(log(dem)))

ggplot(obs_p) +
  geom_histogram(aes(-log(imp)))

ggplot(obs_p) +
  geom_histogram(aes(imp))

ggplot(obs_p) +
  geom_histogram(aes(TN12am), binwidth = 1) +
  xlim(9, 30)
```

```{r}
reg_lm <- aov(tmin ~ TN12am, data = train_rnds)
res_lm <- reg_lm$residuals
fit_lm <- reg_lm$fitted.values
plot_res(fit_lm, res_lm, "ANOVA tmin ~ TN12am")
plot_reg(train_rnds$tmin, fit_lm, "ANOVA tmin ~ TN12am")
summary(reg_lm)
```

```{r}
reg_lm <- aov(formula, data = train_rnds)
res_lm <- reg_lm$residuals
fit_lm <- reg_lm$fitted.values
plot_res(fit_lm, res_lm, "ANCOVA without covariate selection")
plot_reg(
  train_rnds$tmin, fit_lm,
  paste("ANCOVA \n",
        predicted, " ~ \n", 
        paste(predictors, collapse= " + "))
)
summary(reg_lm)
```
Regression shrinkage methods


```{r}
reg_lasso.quanti <- glmnet(
  y = as.data.frame(train_rnds)[, predicted],
  x = as.matrix(
    as.data.frame(train_rnds)[, predictors_q]
  )
)
plot(reg_lasso.quanti, xvar = "lambda", label = TRUE)
legend(
  x = -4, y = 0.2,
  legend = paste(1:length(predictors_q), " - ", predictors_q),
  cex = .6
)
```

```{r}
x_mat <- model.matrix(formula, data = train_rnds)
x_mat <- x_mat[, 2:ncol(x_mat)] # -- remove intercept
reg_lasso <- glmnet(
  y = as.data.frame(train_rnds)[, predicted],
  x = x_mat
)


options(repr.plot.width = 12, repr.plot.height = 10)
plot(reg_lasso, xvar = "lambda", label = TRUE)
legend(
  x = -1, y = -0.1,
  legend = paste(1:ncol(x_mat), " - ", colnames(x_mat)),
  cex = .4
)
```

Optimum for log(lambda) : two covariates (TN12am and dem)

Classical cross validation 

```{r}
reg_lasso_cv <- cv.glmnet(y = as.data.frame(train_rnds)[,predicted], 
                          x = x_mat)
plot(reg_lasso_cv)
```

```{r}
paste("CV estimate of lambda :", round(reg_lasso_cv$lambda.1se, 3))
coef(reg_lasso_cv, s = "lambda.1se")
```

```{r}
fit_lasso_1se <- stats::predict(reg_lasso_cv, s = "lambda.1se", 
                         newx = x_mat)
res_lasso_1se <- data.frame(train_rnds)[, predicted] - fit_lasso_1se 

# Graphe des résidus
plot_res(fit_lm, res_lm, "Linéaire, sans sélection")
plot_res(fit_lasso_1se, res_lasso_1se, 
         "Linéaire penalité L1 lambda 1se") 

plot_reg(data.frame(train_rnds)[, predicted], fit_lm, 
         "Linéaire, sans sélection") 
plot_reg(data.frame(train_rnds)[, predicted], fit_lasso_1se, 
         "Linéaire penalité L1 lambda 1se") 
```

```{r}
paste("Without coef shrinkage MSE:", mean(res_lm^2))
paste("LASSO with lambda.1se:", mean(res_lasso_1se^2))
```

10-fold CV mean square error 
(traditional one, see to implement spatial one)

```{r}
V <- 10
nV <- floor(nrow(train_rnds) / V)
S <- sample(1:nrow(train_rnds), replace = FALSE)
error_CV <- c()
for (v in 1:V) {
  train_rnds_learn <- train_rnds[-c(S[(nV * (v - 1)):(nV * v)]), ]
  train_rnds_valid <- train_rnds[c(S[(nV * (v - 1)):(nV * v)]), ]
  fit_valid <- predict(aov(formula, data = train_rnds_learn), 
                 newdata = train_rnds_valid)
  y_valid <- data.frame(train_rnds_valid)[, predicted]
  error_CV <- c(error_CV, mean((y_valid - fit_valid)^2))
}
mean(error_CV)
print(reg_lasso_cv)
```

Quadratic model (AIC criteria model selection)

```{r}
reg_glm <- glm(tmin ~ .^2, data = train_rnds[, c("tmin", "tcc", "imp", "build.fp", "dem", "TN12am")])


reg_glm_step <- step(reg_glm, direction = "backward")
anova(reg_glm_step, test_rnds = "F")   
```

LASSO (L1 shrinkage)

```{r}
x_mat2 <- model.matrix(tmin ~ .^2 - 1, data = train_rnds[, c("tmin", "tcc", "imp", "build.fp", "dem", "TN12am")])
reg_lasso2 <- glmnet(y = train_rnds$tmin, x = x_mat2)
reg_lasso2_cv <- cv.glmnet(y = train_rnds$tmin, x = x_mat2)
coef(reg_lasso2_cv, s = "lambda.1se")
```

```{r}
fit_glm <- reg_glm_step$fitted.values 
res_glm <- reg_glm_step$residuals
fit_lasso2_cv <- predict(reg_lasso2_cv, s = "lambda.min", newx = x_mat2)
res_lasso2_cv <- train_rnds$tmin - fit_lasso2_cv

# Graphe des résidus
g1<-gplot_res(fit_lm, res_lm, "linéaire")
g2<-gplot_res(fit_lasso_1se, res_lasso_1se, "linéaire, pénalité L1")
g3<-gplot_res(fit_glm, res_glm, "quadratique, backward AIC")
g4<-gplot_res(fit_lasso2_cv, res_lasso2_cv, "quadratique, pénalité L1")
grid.arrange(g1,g2,g3,g4,ncol=2,nrow=2)
```

test_rnds sample predictions

```{r}
test_rnds$glm.pred <- predict(reg_glm_step, newdata = test_rnds)
test_rnds$glm.res <- test_rnds$glm.pred - test_rnds$tmin
sum((test_rnds$glm.res)^2) / nrow(test_rnds)
```

Grid prediction

```{r}
pred_p$glm.pred <- predict(reg_glm_step, newdata = pred_p)
```

Select and plot a specific date

```{r}
d <- "2022-07-20"
plot.pred <- pred_p[pred_p$date==d,] 
plot.pred <- st_as_sf(plot.pred, 
                    coords=c("lon", "lat"),
                    remove=F,
                    crs=4326)
plot.pred <- vect(plot.pred)

m <- ggplot() +
  geom_spatvector(data = plot.pred, aes(color=glm.pred), size=.1) +
  #geom_sf(data = obs_p, aes(geometry=geometry, color=tmin)) +
  geom_sf(data = st_as_sf(nc_borders), aes(geometry=geometry), 
               colour = "grey", linewidth=.3, fill = NA) +
  scale_color_whitebox_c(
      palette = "muted",
      labels = scales::label_number(suffix = "°C"),
      n.breaks = 12,
      guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Generalized Linear Model"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot_caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
m
```

#### Compare results on different definitions of train - test

Results with LASSO regression on quadratic covariates

```{r}
x_mat2 <- model.matrix(tmin ~ .^2 - 1, data = train_rnds[, c("tmin", "tcc", "imp", "build.fp", "dem", "TN12am")])
x_mat2.test <- model.matrix(tmin ~ .^2 - 1, data = test_rnds[, c("tmin", "tcc", "imp", "build.fp", "dem", "TN12am")])

reg_lasso2 <- glmnet(y = train_rnds$tmin, x = x_mat2)
test_rnds$lasso2.pred <- predict(reg_lasso2, newx = x_mat2.test, s=0.1)

test_rnds$lasso2.res <- test_rnds$lasso2.pred - test_rnds$tmin
sum((test_rnds$lasso2.res)^2) / nrow(test_rnds)
pred_p$lasso2.rnds.pred <- predict(reg_lasso2, newdata = pred_p)
```

## Random Forest

```{r}
rf.reg <- randomForest(
  tmin ~ tcc + imp + build.fp + build.h + dem + TN12am, 
  data=train_rnds, 
  xtest_rnds=test_rnds[, .(tcc,imp,build.fp,build.h,dem,TN12am)], 
  ytest_rnds=test_rnds$tmin,
  ntree=500, mtry=2, do.trace=50, importance=TRUE)
attributes(rf.reg)
rf.reg$mtry
```

```{r}
fit_rf=rf.reg$predicted
res_rf=fit_rf-train_rnds$tmin
gplot_res(fit_rf,res_rf,titre="")
```

Variable importance

```{r}
rf.reg$importance
varImpPlot(rf.reg)
library(ggRandomForests)
plot(gg_vimp(rf.reg))
```

On test_rnds sample

```{r}
test_rnds$rf.pred <- predict(rf.reg, newdata = test_rnds)
#test_rnds$rf.pred <- rf.reg$test_rnds$predicted
test_rnds$rf.res <- test_rnds$rf.pred - test_rnds$tmin
sum((test_rnds$rf.res)^2)/nrow(test_rnds)
```

Prediction on the grid

```{r}
pred.rf.grid <- predict(object=rf.reg, newdata = pred_p)
pred_p$rf.pred <- pred.rf.grid 
```

```{r}
d <- "2022-07-20"
plot.pred <- pred_p[pred_p$date==d,] 
plot.pred <- st_as_sf(plot.pred, 
                    coords=c("lon", "lat"),
                    remove=F,
                    crs=4326)
plot.pred <- vect(plot.pred)

m <- ggplot() +
  geom_spatvector(data = plot.pred, aes(color=rf.pred), size=.1) +
  #geom_sf(data = obs_p, aes(geometry=geometry, color=tmin)) +
  geom_sf(data = st_as_sf(nc_borders), aes(geometry=geometry), 
               colour = "grey", linewidth=.3, fill = NA) +
  scale_color_whitebox_c(
      palette = "muted",
      labels = scales::label_number(suffix = "°C"),
      n.breaks = 12,
      guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Generalized Linear Model"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot_caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
m
```

#### Gradient Boosting

```{r}

```

#### Neural networks

```{r}

```

## Kriging on residuals

-   compute predictions on test_rnds set

-   plot empirical variogram of residuals on validation set to see if there is a spatio-temporal pattern

-   fit

#### Variogram analysis

Tranform test_rnds set to a STFDF object from spacetime library

```{r}
library(spacetime)
test_rnds <- test_rnds[, ":="(date=as.Date(date)),]
test_rnds_st <- stConstruct(as.data.frame(test_rnds), space=c("lon","lat"), time="date", 
                       SpatialObj=SpatialPoints(test_rnds[,c("lon","lat")]), 
                       crs=CRS(crs_nad83))
test_rnds_st <- as(test_rnds_st,"STFDF")
```

Plot residulals for one date

Plot residuals evolution per one location

```{r, eval=FALSE}
ggplot(test_rnds[which(test_rnds$station %in% unique(test_rnds$station)[1:10]),])+
  geom_line(aes(x=date, y=glm.res, color=station, group=station))+
  geom_point(aes(x=date, y=glm.res, color=station, group=station))+
  theme(legend.position="none")


ggplot(test_rnds[which(test_rnds$network=="ECONET"),])+
  geom_line(aes(x=date, y=glm.res, color=station, group=station))+
  geom_point(aes(x=date, y=glm.res, color=station, group=station))+
  ylim(-5,5)+
  theme(legend.position="none")

ggplot(test_rnds[which(test_rnds$network=="RAWS"),])+
  geom_line(aes(x=date, y=glm.res, color=station, group=station))+
  geom_point(aes(x=date, y=glm.res, color=station, group=station))+
  ylim(-5,5)+
  theme(legend.position="none")

ggplot(test_rnds[which(test_rnds$network=="WBAN"),])+
  geom_line(aes(x=date, y=glm.res, color=station, group=station))+
  geom_point(aes(x=date, y=glm.res, color=station, group=station))+
  ylim(-5,5)+
  theme(legend.position="none")

ggplot(test_rnds[which(test_rnds$network=="COOP"),])+
  geom_line(aes(x=date, y=glm.res, color=station, group=station))+
  geom_point(aes(x=date, y=glm.res, color=station, group=station))+
  ylim(-5,5)+
  theme(legend.position="none")

ggplot(test_rnds)+
  geom_line(aes(x=date, y=glm.res, color=station, group=station))+
  geom_point(aes(x=date, y=glm.res, color=station, group=station))+
  facet_wrap(vars(network))+
  ylim(-5,5)+
  theme(legend.position="none")


ggplot(test_rnds)+
  geom_point(aes(x=lon, y=lat, color=glm.res))+
  geom_point(data=train_rnds, aes(x=lon, y=lat), shape=3, color="grey")+
  facet_wrap(vars(date))+
  scale_color_steps2(low="blue", mid="white", high="red", limits=c(-3,3), n.breaks=10)

```

Directions: 0=north, 90=east, 180=south, 270=west

```{r, eval=F}
cutoff=30000 
width=1000
plot(variogram(rf.res ~ 1, data=test_rnds_st, cloud = TRUE, cutoff=cutoff, width=width), tlags=0:3, main="semivariogram cloud residuals RF")


plot(variogram(glm.res ~ 1, test_rnds_st, cloud = TRUE, cutoff=10000, width=width), main="semivariogram cloud residuals glm")

plot(variogram(tmin ~ 1, obs.d, cutoff=cutoff, width=width), main="sample semivariogram")
plot(variogram(tmin ~ 1, obs.d, cutoff=cutoff, width=width, alpha = c(0, 45, 90, 135, 180, 225, 270, 315)), main="direction dependance")

```

Variogram model

```{r, eval=F}
cutoff <- 100000
v.emp <- variogram(tmin ~ 1, obs.d, cutoff=cutoff, width=width)

nugget <- 0.5
range <- 20000
vgm.sph <- vgm(psill=3, "Sph", range=range, nugget=0.5, cutoff=cutoff)
plot(v.emp, vgm.sph, cutoff=cutoff)

vgm.exp <- vgm(psill=3, "Exp", range=range, nugget=0.5, cutoff=cutoff)
plot(v.emp, vgm.exp, cutoff=cutoff)

vgm.mat <- vgm(psill=3, "Mat", range=range, nugget=0.5, cutoff=cutoff)
plot(v.emp, vgm.mat, cutoff=cutoff)
```

Fit Matern model

```{r, eval=F}
v.fit <- fit.variogram(v.emp, vgm.mat, fit_range=F)
plot(v.emp, v.fit)
```

#### Ordinary kriging on residuals

```{r, eval=F}
cutoff <- 100000
nugget <- 0.5
range <- 20000
width <- 1000
v.emp <- variogram(tmin ~ 1, obs.d, cutoff=cutoff, width=width)
vgm.mat <- vgm(psill=3, "Mat", range=range, nugget=nugget, cutoff=cutoff)
v.fit <- fit.variogram(v.emp, vgm.mat, fit_range=F)
ok <- gstat::krige(tmin ~ 1, obs.d, newdata=pred.d, v.fit)
```

```{r, eval=F}
m <- ggplot() +
  geom_spatvector(data = vect(ok), aes(color=var1.pred), size=.1) +
  geom_sf(data = obs.d, aes(geometry=geometry, color=tmin)) +
  geom_sf(data = st_as_sf(nc_borders), aes(geometry=geometry), 
               colour = "grey", linewidth=.3, fill = NA) +
  scale_color_whitebox_c(
      palette = "muted",
      labels = scales::label_number(suffix = "°C"),
      n.breaks = 12,
      guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Ordinary kriging"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot_caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
m
```
