---
title: "spatiotemporal-variography"
output: html_document
date: "2023-08-23"
---

Libraries

```{r, message=F}
library(terra)
library(sf)

library(ggplot2)
library(ggspatial)
library(tidyterra)
library(rgeos)
library(data.table) # -- for large flat datasets

# space-time variography
library(gstat)
```

```{r}
input_path <- "../input/"
```

## Open observations

```{r}
aws_file <- paste0("NC-AWS-NOAA-dailysummary-20220601-20220831",
                   "-space-time-covariates.csv")
aws <- fread(paste0(input_path, aws_file))
```

#### Plot TMIN for each month

```{r}
ggplot(aws[between(DATE, as.Date("2022-05-31"), as.Date("2022-06-30"))]) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE, color = TMIN)) +
  facet_wrap(~DATE) +
  scale_color_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "°C"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  coord_equal()

ggsave("../output/noaa_obs_202206_TMIN.png", width = 15, height = 5, dpi = 300)

ggplot(aws[between(DATE, as.Date("2022-07-01"), as.Date("2022-07-31"))]) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE, color = TMIN)) +
  facet_wrap(~DATE) +
  scale_color_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "°C"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  coord_equal()

ggsave("../output/noaa_obs_202207_TMIN.png", width = 15, height = 6, dpi = 300)

ggplot(aws[between(DATE, as.Date("2022-08-01"), as.Date("2022-08-31"))]) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE, color = TMIN)) +
  facet_wrap(~DATE) +
  scale_color_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "°C"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  coord_equal()

ggsave("../output/noaa_obs_202208_TMIN.png", width = 15, height = 6, dpi = 300)
```

#### Plot TMAX for each month

```{r}
ggplot(aws[between(DATE, as.Date("2022-05-31"), as.Date("2022-06-30"))]) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE, color = TMAX)) +
  facet_wrap(~DATE) +
  scale_color_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "°C"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  coord_equal()

ggsave("../output/noaa_obs_202206_TMAX.png", width = 15, height = 5, dpi = 300)

ggplot(aws[between(DATE, as.Date("2022-07-01"), as.Date("2022-07-31"))]) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE, color = TMAX)) +
  facet_wrap(~DATE) +
  scale_color_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "°C"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  coord_equal()

ggsave("../output/noaa_obs_202207_TMAX.png", width = 15, height = 6, dpi = 300)

ggplot(aws[between(DATE, as.Date("2022-08-01"), as.Date("2022-08-31"))]) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE, color = TMAX)) +
  facet_wrap(~DATE) +
  scale_color_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "°C"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  coord_equal()

ggsave("../output/noaa_obs_202208_TMAX.png", width = 15, height = 6, dpi = 300)
```

```{r}
ggplot(aws) +
  geom_tile(aes(x = as.factor(LONGITUDE), y = DATE, fill = TMIN)) +
  scale_y_date(
    date_labels = "%d/%m", date_breaks = "7 days",
    date_minor_breaks = "1 day",
    limits = c(as.Date("2022-06-01", tz = "UTC"),
               as.Date("2022-08-31", tz = "UTC"))
  ) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "°C"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  ggtitle("TMIN") +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 16),
    axis.title.x = element_text(size = 22),
    axis.title.y = element_text(size = 22),
    legend.key.width = unit(1, "cm"),
    panel.grid.major = element_line(color = "grey", size = 0.2),
    legend.text = element_text(size = 16),
    plot.caption = element_text(size = 14),
    legend.title = element_text(size = 18),
    legend.text.align = 0,
    legend.box.spacing = unit(0, "pt")
  )

ggplot(aws) +
  geom_tile(aes(x = as.factor(LONGITUDE), y = DATE, fill = TMAX)) +
  scale_y_date(
    date_labels = "%d/%m", date_breaks = "7 days",
    date_minor_breaks = "1 day",
    limits = c(as.Date("2022-06-01", tz = "UTC"),
               as.Date("2022-08-31", tz = "UTC"))
  ) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "°C"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  ggtitle("TMAX") +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 16),
    axis.title.x = element_text(size = 22),
    axis.title.y = element_text(size = 22),
    legend.key.width = unit(1, "cm"),
    panel.grid.major = element_line(color = "grey", size = 0.2),
    legend.text = element_text(size = 16),
    plot.caption = element_text(size = 14),
    legend.title = element_text(size = 18),
    legend.text.align = 0,
    legend.box.spacing = unit(0, "pt")
  )
```

TMIN and TMAX according to covariates

```{r}
ggplot(aws) +
  geom_point(aes(x = LONGITUDE, y = TMIN), alpha = .2)

ggplot(aws) +
  geom_point(aes(x = LATITUDE, y = TMIN), alpha = .2)

ggplot(aws) +
  geom_point(aes(x = dem, y = TMIN), alpha = .2)

ggplot(aws) +
  geom_point(aes(x = tcc, y = TMIN), alpha = .2)

ggplot(aws) +
  geom_point(aes(x = build.fp, y = TMIN), alpha = .2)

ggplot(aws) +
  geom_point(aes(x = imp, y = TMIN), alpha = .2)

ggplot(aws) +
  geom_boxplot(aes(group = build.h, x = build.h, y = TMIN))
```

```{r}
ggplot(aws) +
  geom_point(aes(x = LONGITUDE, y = TMAX), alpha = .2)

ggplot(aws) +
  geom_point(aes(x = LATITUDE, y = TMAX), alpha = .2)

ggplot(aws) +
  geom_point(aes(x = dem, y = TMAX), alpha = .2)

ggplot(aws) +
  geom_point(aes(x = tcc, y = TMAX), alpha = .2)

ggplot(aws) +
  geom_point(aes(x = build.fp, y = TMAX), alpha = .2)

ggplot(aws) +
  geom_point(aes(x = imp, y = TMAX), alpha = .2)

ggplot(aws) +
  geom_boxplot(aes(group = build.h, x = build.h, y = TMAX))
```

```{r}
summary.jja <- aws[, .(
  TMIN.mean = mean(TMIN),
  TMAX.mean = mean(TMAX),
  TMIN.min = min(TMIN),
  TMAX.min = min(TMAX),
  TMIN.max = max(TMIN),
  TMAX.max = max(TMAX)
), keyby = c("DATE")]
```

```{r}
ggplot() +
  geom_line(data = aws, aes(y = TMIN, x = DATE, group = STATION),
            color = "navyblue", alpha = 0.1) +
  geom_line(data = summary.jja, aes(y = TMIN.mean, x = DATE),
            color = "black", alpha = 1, linewidth = .7) +
  geom_line(data = summary.jja, aes(y = TMIN.min, x = DATE),
            color = "black", alpha = 1, linewidth = .7, linetype = "dashed") +
  geom_hline(yintercept = 20, color = "orange") +
  geom_line(data = summary.jja, aes(y = TMIN.max, x = DATE),
            color = "black", alpha = 1, linewidth = .7, linetype = "dashed") +
  theme_minimal()
ggsave("../output/noaa_obs_jja_timeseries_TMIN.png",
       width = 15, height = 5, dpi = 300)

ggplot() +
  geom_line(data = aws, aes(y = TMAX, x = DATE, group = STATION),
            color = "red", alpha = 0.1) +
  geom_line(data = summary.jja, aes(y = TMAX.mean, x = DATE),
            color = "black", alpha = 1, linewidth = .7) +
  geom_line(data = summary.jja, aes(y = TMAX.min, x = DATE),
            color = "black", alpha = 1, linewidth = .7, linetype = "dashed") +
  geom_hline(yintercept = 35, color = "orange") +
  geom_line(data = summary.jja, aes(y = TMAX.max, x = DATE),
            color = "black", alpha = 1, linewidth = .7, linetype = "dashed") +
  theme_minimal()

ggsave("../output/noaa_obs_jja_timeseries_TMAX.png",
       width = 15, height = 5, dpi = 300)
```

#### Space-time covariography

Create spacetime object from aws

```{r}
library(spacetime)
aws <- aws[, ":="(DATE = as.Date(DATE)), ]
coord <- aws[, c("LONGITUDE", "LATITUDE")]
aws_st <- stConstruct(aws,
                      space = c("LONGITUDE", "LATITUDE"),
                      time = "DATE",
                      SpatialObj = SpatialPoints(coord),
                      crs = crs("EPSG:4326"))
aws_st <- as(aws_st, "STFDF")
```

Spatial autocorrelation for one single date

```{r}
library(gstat)
date <- "2022-08-03"

v <- variogram(TMIN ~ 1,
  aws_st[
    !is.na(aws_st[, date, "TMIN"]$TMIN),
    date, "TMIN"
  ],
  cutoff = 20,
  width = 1
)
plot(v, main = "TMIN")

v <- variogram(tcc ~ 1,
  aws_st[
    !is.na(aws_st[, date, "tcc"]$tcc),
    date, "tcc"
  ],
  cutoff = 20,
  width = 1
)
plot(v, main = "tcc")

v <- variogram(dem ~ 1,
  aws_st[
    !is.na(aws_st[, date, "dem"]$dem),
    date, "dem"
  ],
  cutoff = 30,
  width = 2
)
plot(v, main = "dem")
```

Mean variogram on multiple dates

```{r}
lst <- lapply(1:90, function(i) {
  x <- aws_st[, i, "TMIN"]
  x$ti <- i
  rownames(x@coords) <- NULL
  x
})
pts <- do.call(rbind, lst)
vv <- variogram(TMIN ~ ti,
  pts[!is.na(pts$TMIN), ],
  cutoff = 50,
  width = 2
)
plot(vv, main = "TMIN")
```

Space-time semi-variogram

```{r}
vvst <- variogram(TMIN ~ 1,
  aws_st[!is.na(aws_st[, , "TMIN"]$TMIN), , "TMIN"],
  width = 5,
  cutoff = 80,
  tlags = 0:8
)
plot(vvst, main = "TMIN")
plot(vvst, main = "TMIN", map = FALSE)

vvst <- variogram(TMAX ~ 1,
  aws_st[!is.na(aws_st[, , "TMAX"]$TMAX), , "TMAX"],
  width = 5,
  cutoff = 80,
  tlags = 0:8
)
plot(vvst, main = "TMAX")
plot(vvst, main = "TMAX", map = FALSE)
```

Covariates colinearities

```{r}
ggplot(aws[DATE == "2022-06-02"]) +
  geom_point(aes(x = tcc, y = imp), alpha = 0.5)

ggplot(aws[DATE == "2022-06-02"]) +
  geom_point(aes(x = tcc, y = build.fp / 9), alpha = 0.5)

ggplot(aws[DATE == "2022-06-02"]) +
  geom_point(aes(x = imp, y = build.fp / 9), alpha = 0.5)

ggplot(aws[DATE == "2022-06-02"]) +
  geom_point(aes(x = tcc, y = dem), alpha = 0.5)

ggplot(aws[DATE == "2022-06-02"]) +
  geom_point(aes(x = imp, y = dem), alpha = 0.5)
```

```{r}
ggplot(aws[DATE == "2022-06-02"]) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE, color = dem)) +
  scale_color_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "m"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  coord_equal()

ggplot(aws[DATE == "2022-06-02"]) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE, color = tcc)) +
  scale_color_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "%"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  coord_equal()

ggplot(aws[DATE == "2022-06-02"]) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE, color = imp)) +
  scale_color_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "%"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  coord_equal()

ggplot(aws[DATE == "2022-06-02"]) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE, color = build.fp / 900 * 100)) +
  scale_color_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "%"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  coord_equal()

ggplot(aws[DATE == "2022-06-02"]) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE, color = build.h)) +
  coord_equal()
```

PCA

```{r}
library(factoextra)
res_pca <- prcomp(aws[DATE == "2022-06-02",
                      .(dem, imp, tcc, build.fp, TX12am, TN12am)],
                  scale = TRUE)
fviz_eig(res_pca)

fviz_pca_ind(res_pca, geom = "point", alpha = 0.5)
fviz_pca_var(res_pca,
  col.var = "contrib", # Color by contributions to the PC
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  repel = TRUE # Avoid text overlapping
)

fviz_pca_biplot(res_pca,
  repel = TRUE,
  col.var = "#2E9FDF", # Variables color
  col.ind = "#696969", # Individuals color
  geom = "point"
)
```

```{r}
res_pca <- prcomp(aws[DATE == "2022-06-20",
                      .(dem, imp, tcc, build.fp, TX12am, TN12am)],
                  scale = TRUE)
fviz_eig(res_pca)

fviz_pca_ind(res_pca, geom = "point", alpha = 0.5)
fviz_pca_var(res_pca,
  col.var = "contrib", # Color by contributions to the PC
  gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
  repel = TRUE # Avoid text overlapping
)
```

Kmeans on prediction grid

```{r}
library(terra)
library(data.table)
grid_spat <- vect("../input/prediction_grid_points_urb_rur_spatialcov.shp")
grid_spat$lon <- geom(grid_spat)[, "x"]
grid_spat$lat <- geom(grid_spat)[, "y"]
grid_spat <- setDT(as.data.frame(grid_spat))
dim(grid_spat)
```

Crops and water bodies are set to tcc==0 and imp==0

Shouldn't we create another variable such as 'degree of vegetalization' instead of tcc? Or height of vegetation?

```{r}
dt_plot <- grid_spat[, .(imp = as.numeric(imp), tcc = as.numeric(tcc),
                         dem = as.numeric(dem), build.fp = as.numeric(build.fp),
                         build.h = build.h, lon = as.numeric(lon),
                         lat = as.numeric(lat)), ]

ggplot(dt_plot[imp == 0 & tcc == 0, , ]) +
  geom_point(aes(x = lon, y = lat), size = 0.01) +
  coord_equal()
```

```{r}
library(factoextra)
set.seed(123)

data_cluster <- dt_plot[, .(imp, build.fp / 9, tcc, dem)]
data_cluster <- scale(data_cluster)

# nb clusters
fviz_nbclust(data_cluster[1:20000, ], kmeans, method = "silhouette")

res_km <- kmeans(data_cluster, 5)
fviz_cluster(res_km,
  data = data_cluster,
  palette = c("#2E9FDF", "#00AFBB", "#E7B800", "orange", "red"),
  geom = "point",
  ellipse.type = "convex",
  ggtheme = theme_bw(),
  alpha = 0.1
)

dt_plot$cluster <- res_km$cluster
ggplot(dt_plot) +
  geom_point(aes(x = lon, y = lat, color = as.factor(cluster)), size = 0.01) +
  scale_color_manual(values = c("#2E9FDF", "yellow", "green",
                                "purple", "red")) +
  coord_equal()
```
