---
title: "spatial_lm_temperature"
output: html_document
date: "2023-10-20"
---


Libraries

```{r, message=F}
library(gstat)
library(terra)
library(sf)
library(sftime)
library(tidyverse)
library(ggplot2)
library(ggspatial)
library(gridExtra)
library(tidyterra)
library(rgeos)
library(data.table) # -- for large flat datasets
library(DT)

# -- for timeseries
library(lubridate)
library(xts)

# -- ml libraries
library(caret)
library(glmnet)
library(randomForest)

# -- code style
library(styler)

# -- spatial k-fold CV
library(spatialsample)

# -- tidy tools for quantifying how well model fits to a data set
library(yardstick)
```

My own functions to manipulate data:

```{r}
source("../R/manipulate_spacetime_data.R")
source("../R/open_data_heatmodel.R")
source("../R/ml_check_input_data.R")
source("../R/ml_create_spatial_train_test_sets.R")
source("../R/ml_create_spatial_cross_validation.R")
source("../R/ml_define_models.R")
source("../R/ml_compute_plot_scores_outputs.R")
```

#### Open data on a specific period

```{r}
obs <- fread("../input/NC-monitors-dailysummary-20220601-20220831-space-time-covariates.csv")
nc_borders <- vect("../input/NC_county_boundary/North_Carolina_State_and_County_Boundary_Polygons.shp")

p <- seq(as.Date("2022-07-20"), as.Date("2022-07-22"), by = "1 day")
pred_p <- open_pred_period(p)
obs_p <- obs[date %in% p, ]
```

Define origin and destination projections

```{r}
crs_wgs84 <- "epsg:4326"
crs_nad83 <- "epsg:5070"
```

Project data.tables lon and lat variables

```{r}
pred_p <- project_dt(pred_p, crs_wgs84, crs_nad83)
obs_p <- project_dt(obs_p, crs_wgs84, crs_nad83)
```

Add county to data.tables

```{r}
pred_p <- add_nc_county(pred_p, crs_nad83)
obs_p <- add_nc_county(obs_p, crs_nad83)
```

Remove NA in obs_p_sft

```{r}
obs_p <- obs_p[!is.na(obs_p$tmin) & !is.na(obs_p$tmax), ]
```

#### Check data content

```{r}
meta_obs <- c("lon", "lat", "date", "station", "county")
meta_pred <- c("lon", "lat", "date")
predictors <- c("tcc", "imp", "build.fp", "build.h", "dem", "TN12am")
# -- quantitative predictors
predictors_q <- c("tcc", "imp", "build.fp", "dem", "TN12am")
predicted <- "tmin"
check_obs(obs_p, meta_obs, predictors, predicted)
check_pred_grid(pred_p, meta_pred, predictors)
```


#### Prepare train and test samples

Create several types of train and test sets:

-   rndst: select randomly the observations in space and time

-   rnds: select randomly the stations (and keep the whole timeserie)

-   t: seperate train and test by dates (here: remove the last date of the dataset to evaluate capacity of prediction)

-   [area_name]: seperate train and test by selecting stations (and their entire timeserie) according to a contiguous area (group of counties). Several options are possible (test can be urban - mountain - piedmont - sea - plain).

-   net: select monitors per network. Reliable networks: ECONET, CRN, WBAN. Less reliable: COOP and RAWS.

```{r}
set.seed(111)

# -- regular 1-fold
sets_rndst <- create_sets_rndst(obs_p)
train_rndst <- sets_rndst$train
test_rndst <- sets_rndst$test

# -- randomly 1-fold locations out
sets_rnds <- create_sets_rnds(obs_p)
train_rnds <- sets_rnds$train
test_rnds <- sets_rnds$test

# -- randomly 1-fold dates out
test_dates <- as.Date("2022-07-22")
sets_t <- create_sets_t(obs_p, test_dates)
train_t <- sets_t$train
test_t <- sets_t$test

# -- blocks 1-fold locations out

urb <- c("Durham", "Orange", "Chatham", "Wake")
sets_s <- create_sets_s(obs_p, urb)
train_urb <- sets_s$train
test_urb <- sets_s$test

mou <- c("Wilkes", "Ashe", "Watauga", "Caldwell")
sets_s <- create_sets_s(obs_p, mou)
train_mou <- sets_s$train
test_mou <- sets_s$test

pie <- c("Rowan", "Davidson", "Davie", "Forsyth")
sets_s <- create_sets_s(obs_p, pie)
train_pie <- sets_s$train
test_pie <- sets_s$test

pla <- c("Wilson", "Edgecombe", "Nash", "Halifax")
sets_s <- create_sets_s(obs_p, pla)
train_pla <- sets_s$train
test_pla <- sets_s$test

sea <- c("Tyrrell", "Washington", "Dare", "Hyde")
sets_s <- create_sets_s(obs_p, sea)
train_sea <- sets_s$train
test_sea <- sets_s$test

# -- selection with network origin
net <- c("ECONET", "CRN", "WBAN")
sets_net <- create_sets_net(obs_p, net)
train_net <- sets_net$train
test_net <- sets_net$test
```

Plot training et testing sample tile plots and maps

```{r}
df_test <- test_rndst
df_test$set <- "test"
df_train <- train_rndst
df_train$set <- "train"
df <- rbind(df_train, df_test)

ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile() +
  labs(title = "Test on random observations") +
  scale_fill_manual(values = c(train = "lightblue", 
                               test = "tomato4")) +
  my_theme_paper() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
```

```{r}
df_test <- test_rnds
df_test$set <- "test"
df_train <- train_rnds
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile() +
  labs(title = "Test on random monitors") +
  scale_fill_manual(values = c(train = "lightblue", 
                               test = "tomato4")) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
```

```{r}
df_test <- test_t
df_test$set <- "test"
df_train <- train_t
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile() +
  labs(title = "Test on last day") +
  scale_fill_manual(values = c(train = "lightblue", 
                               test = "tomato4")) +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
```

```{r}
df_test <- test_urb
df_test$set <- "test"
df_train <- train_urb
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile() +
  scale_fill_manual(values = c("train" = "lightblue", test = "tomato4")) +
  labs(title = "Test on urban county") +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
ggplot(df, aes(x = lon, y = lat, color = set)) +
  geom_point() +
  scale_color_manual(values = c("train" = "lightblue", test = "tomato4")) +
  labs(title = "Test on urban counties") +
  coord_equal()
```

```{r}
df_test <- test_mou
df_test$set <- "test"
df_train <- train_mou
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile() +
  scale_fill_manual(values = c("train" = "lightblue", test = "tomato4")) +
  labs(title = "Test on mountain counties") +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  )
ggplot(df, aes(x = lon, y = lat, color = set)) +
  geom_point() +
  scale_color_manual(values = c("train" = "lightblue", test = "tomato4")) +
  labs(title = "Test on mountain counties") +
  coord_equal()
```

```{r}
df_test <- test_pie
df_test$set <- "test"
df_train <- train_pie
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile()+
  scale_fill_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on Piedmont counties")+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()
    )  
ggplot(df, aes(x = lon, y = lat, color = set)) +
  geom_point()+
  scale_color_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on Piedmont counties")+
  coord_equal()
```

```{r}
df_test <- test_pla
df_test$set <- "test"
df_train <- train_pla
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile()+
  scale_fill_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on plain counties")+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()
    )  
ggplot(df, aes(x = lon, y = lat, color = set)) +
  geom_point()+
  scale_color_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on plain counties")+
  coord_equal()
```

```{r}
df_test <- test_sea
df_test$set <- "test"
df_train <- train_sea
df_train$set <- "train"
df <- rbind(df_train, df_test)
ggplot(df, aes(y = station, x = as.Date(date), fill = set)) +
  geom_tile()+
  scale_fill_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on sea counties")+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank()
    )  
ggplot(df, aes(x = lon, y = lat, color = set)) +
  geom_point()+
  scale_color_manual(values=c("train"="lightblue", test="tomato4"))+
  labs(title="Test on sea counties")+
  coord_equal()
```

#### Prepare spatial k-fold CV

```{r}
obs_p_sf <- create_sf(obs_p, "epsg:5070")
sp_samples <- create_sp_fold(obs_p_sf)
```

Plots 

```{r}
cluster_llo <- spatial_clustering_cv(obs_p_sf, v = 10)
block_llo <- spatial_block_cv(obs_p_sf, v = 10)
net_llo <- spatial_leave_location_out_cv(obs_p_sf,
  group = network,
  v = 6
)
rnd_llo <- spatial_leave_location_out_cv(obs_p_sf,
  group = station,
  v = 10
)
autoplot(cluster_llo)
autoplot(block_llo)
autoplot(net_llo)
autoplot(rnd_llo)
```

#### Fit cross validation 

Formula definition

```{r}
(formula <- as.formula(paste(predicted, " ~ ", 
                             paste(predictors, collapse= "+"))))
(formula_q <- as.formula(paste(predicted, " ~ ", 
                             paste(predictors_q,
                                   collapse= "+"))))
```

```{r}
cv_fit <- compute_preds_lm_cv(sp_samples)
cv_rmse <- compute_rmse_cv(cv_fit, predicted=tmin)
cv_rmse
```

Plot RMSE per fold

```{r}
map_rmse_cv(cv_fit, cv_rmse)
```


```{r}
map_res_cv(cv_fit, tmin)
```

```{r}
plot_reg_cv(cv_fit, tmin)
```
