---
title: "era5-reanalysis-processing"
output: github_document
 #pdf_document: default
#  html_document: default
date: "2023-07-24"
---

Librariesƒ

```{r, message=F}
# -- for spatial data
#library(raster)   # -- old
#library(rgdal)    # -- deprecated in 10/2023
#library(rgeos)    # -- deprecated in 10/2023
#library(maptools) # -- deprecated in 10/2023
options("sp_evolution_status" = 2) # use sf instead of rgdal and rgeos in sp
library(sp)
library(terra)
library(sf)
library(ncdf4)

# -- for timeseries
library(lubridate)
library(zoo)
library(xts)

# -- 
library(ggplot2)
library(ggspatial)
library(tidyterra)
library(maditr)
library(tidyverse)
library(viridis)
```

#### NC shapefile

```{r, warning=FALSE}
nc.borders <- vect("../input/NC_county_boundary/North_Carolina_State_and_County_Boundary_Polygons.shp")
```

Create a function to map rasters with NC borders

```{r}

my.map.raster <- function(raster, date, title){
	plot <- ggplot() +
  geom_spatraster(data = raster) +
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "white", linewidth=.3, fill = NA) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "º"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = title,
    subtitle = date
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
	
  return(plot)
}


my.map.raster.points <- function(raster, points, points.fill, date, title){
  fill <- enquo(points.fill)
	plot <- ggplot() +
  geom_spatraster(data = raster) +
	geom_sf(data = points, 
             aes(geometry=geometry, fill=!!fill), color='black',
             size=2, shape=21)+
	geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "white", linewidth=.3, fill = NA) +
  scale_fill_whitebox_c(
    palette = "muted",
    labels = scales::label_number(suffix = "º"),
    n.breaks = 12,
    guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = title,
    subtitle = date
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
	
  return(plot)
}

```

#### Temperature ERA5 reanalysis

Open netcdf data with all JJA hourly ERA5 reanalysis

```{r}
t2m.rea <- nc_open("../input/era5_hourly_reanalysis_20220601_20220831.nc")
print(t2m.rea)
```

Extract data in vectors and 3d-matrix

```{r}
lon <- ncvar_get(t2m.rea, "longitude")
lat <- ncvar_get(t2m.rea, "latitude")
time <- ncvar_get(t2m.rea, "time")
t2m <- ncvar_get(t2m.rea, "t2m") - 273.15
na.value <- ncatt_get(t2m.rea, "t2m", "_FillValue")$value
t2m[t2m==na.value] <- NA
cat(paste('time units:', ncatt_get(t2m.rea, "time", "units")$value))
time <- as.POSIXct(time*3600, origin="1900-01-01", tz="UTC")
```

Map T2M for one hour

```{r}
# one time
ts <- as.POSIXct("2022-08-02 18:00:00 UTC", tz='UTC')
samp <- t2m[,,which(time==ts)]
# need to add 0.25/2 to 
samp.rast <- rast(t(samp), extent=c(-85.125,-74.875,33.375,37.125), crs='+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0')

my.map.raster(raster=samp.rast, date=paste(ts, 'UTC'), title="T2m North Carolina")
```

Code to create a dataframe from the netcdf

```{r}
# -- /!\ latitude vector is decreasing
lon.lat.time <- as.matrix(expand.grid(lon, rev(lat), time))
t2m.vect <- as.vector(t2m[,ncol(t2m):1,])
t2m.df <- data.frame(cbind(lon.lat.time, t2m.vect))
colnames(t2m.df) <- c("lon", "lat", "time", 't2m')
t2m.df$t2m <- as.numeric(t2m)
t2m.df$time <- as.POSIXct(t2m.df$time, tz="UTC")
pixels <- unique(t2m.df[,c('lon', 'lat')])
pixels$geom <- index(pixels)
t2m.df <- merge(t2m.df, pixels, by=c('lon','lat'))
```

Compute of each day

```{r}
ts <- as.POSIXct("2022-06-01 18:00:00 UTC", tz='UTC')
te <- as.POSIXct("2022-08-31 18:00:00 UTC", tz='UTC')
t2m.ts <- t2m.df[which(between(t2m.df$time, ts, te)), ]

dateTN <- function(x){
  return(ifelse(hour(x)>=18, substr(x+days(1),0,10), substr(x,0,10)))
}

t2m.ts$dateTN <- lapply(t2m.ts$time, dateTN)
t2m.ts$dateTN <- as.character(t2m.ts$dateTN)
summary(t2m.ts[,c('t2m','dateTN', 'geom')])

# transform data.frame to data.table to faster computations
t2m.ts <- setDT(t2m.ts) 
t2m.TN <- t2m.ts[, min(t2m), keyby=c('geom', 'dateTN', 'lon', 'lat')]
colnames(t2m.TN) <- c("geom", "dateTN", "lon", "lat", "t2m")
summary(t2m.TN$t2m)
t2m.TN$lon <- as.numeric(t2m.TN$lon)
t2m.TN$lat <- as.numeric(t2m.TN$lat)

beepr::beep(1)
```

Compute TX of each day

```{r}
ts <- as.POSIXct("2022-06-01 06:00:00 UTC", tz='UTC')
te <- as.POSIXct("2022-08-31 06:00:00 UTC", tz='UTC')
t2m.ts <- t2m.df[which(between(t2m.df$time, ts, te)), ]

dateTX <- function(x){
  return(ifelse(hour(x)<6, substr(x-days(1),0,10), substr(x,0,10)))
}

t2m.ts$dateTX <- lapply(t2m.ts$time, dateTX)
t2m.ts$dateTX <- as.character(t2m.ts$dateTX)
summary(t2m.ts[,c('t2m','dateTX', 'geom')])

# transform data.frame to data.table to faster computations
t2m.ts <- setDT(t2m.ts) 
t2m.TX <- t2m.ts[, max(t2m), keyby=c('geom', 'dateTX', 'lon', 'lat')]
colnames(t2m.TX) <- c("geom", "dateTX", "lon", "lat", "t2m")
summary(t2m.TX$t2m)
t2m.TX$lon <- as.numeric(t2m.TX$lon)
t2m.TX$lat <- as.numeric(t2m.TX$lat)

beepr::beep(1)
```

Merge TN and TM in the same data.frame

```{r}
colnames(t2m.TN) <- c("geom", "date", "lon", "lat", "TN")
colnames(t2m.TX) <- c("geom", "date", "lon", "lat", "TX")
era5.t2m <- merge(t2m.TN, t2m.TX, by=c("geom", "date", "lon", "lat"))
```

### Check coherence with NOAA observations

Open NOAA obs

```{r}
aws <- data.table::fread("../input/NC-AWS-NOAA-dailysummary-20220601-20220831.csv") 
length(unique(aws$STATION))
aws <- aws[which(!(is.na(aws$TMAX)|is.na(aws$TMIN))),] 
aws$DATE <- as.Date(aws$DATE, format = "%Y-%m-%d")
length(unique(aws$STATION))
```

Daily max temperatures

```{r, message=FALSE}
aws.ts.tx <- maditr::dcast(aws[, c('DATE', 'STATION', 'TMAX')], DATE ~ STATION) %>%
  as.xts() 
```

Daily min temperatures

```{r, message=FALSE}
aws.ts.tn <- maditr::dcast(aws[, c('DATE', 'STATION', 'TMIN')], DATE ~ STATION) %>%
  as.xts() 
```

Daily mean temperatures

```{r, message=FALSE}
aws.ts.tm <- maditr::dcast(aws[, c('DATE', 'STATION', 'TAVG')], DATE ~ STATION) %>%
  as.xts() 
```

Timeseries plots

```{r}
plot((aws.ts.tn-32)*5/9)
plot((aws.ts.tx-32)*5/9)
plot((aws.ts.tm-32)*5/9)

boxplot(t((aws.ts.tn-32)*5/9))
boxplot(t((aws.ts.tx-32)*5/9))
boxplot(t((aws.ts.tm-32)*5/9))
```

NOAA aws with less than 5% missing data

```{r , message=FALSE}
nb.na.tn <- lapply(aws.ts.tn, FUN=function(x) sum(is.na(x)))  
nb.na.tn <- as.data.frame(do.call(rbind, nb.na.tn))
nb.na.tn <- cbind(STATION = rownames(nb.na.tn), nb.na.tn)
rownames(nb.na.tn) <- 1:nrow(nb.na.tn)
names(nb.na.tn)[names(nb.na.tn) == 'V1'] <- 'nb.na.tn'

nb.na.tx <- lapply(aws.ts.tx, FUN=function(x) sum(is.na(x)))  
nb.na.tx <- as.data.frame(do.call(rbind, nb.na.tx))
nb.na.tx <- cbind(STATION = rownames(nb.na.tx), nb.na.tx)
rownames(nb.na.tx) <- 1:nrow(nb.na.tx)
names(nb.na.tx)[names(nb.na.tx) == 'V1'] <- 'nb.na.tx'

nb.na.tm <- lapply(aws.ts.tm, FUN=function(x) sum(is.na(x)))  
nb.na.tm <- as.data.frame(do.call(rbind, nb.na.tm))
nb.na.tm <- cbind(STATION = rownames(nb.na.tm), nb.na.tm)
rownames(nb.na.tm) <- 1:nrow(nb.na.tm)
names(nb.na.tm)[names(nb.na.tm) == 'V1'] <- 'nb.na.tm'

stations <- unique(aws[,c('STATION', 'NAME', 'LATITUDE', 'LONGITUDE', 'ELEVATION')])
stations <- list(stations, nb.na.tx, nb.na.tn, nb.na.tm) %>% reduce(full_join, by='STATION')
aws <- list(aws, nb.na.tx, nb.na.tn, nb.na.tm) %>% reduce(full_join, by='STATION')

# -- turn into a sf object to reproject CRS
stations <- st_as_sf(stations, coords=c('LONGITUDE', 'LATITUDE'))
st_crs(stations) <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
stations <- st_transform(stations, crs(nc.borders))
stations <- data.frame(stations)
```

Change aws dataframe projection

```{r, message=F}
# -- turn into a sf object to reproject CRS
aws <- st_as_sf(aws, coords=c('LONGITUDE', 'LATITUDE'))
st_crs(aws) <- '+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0'
aws <- st_transform(aws, crs(nc.borders))
aws <- data.frame(aws)
```

Function to extract a date in era5 TX and TN data.table

```{r}
era5.date <- function(date, era5){
  era5.date <- era5[which(era5$date==date), ]
  era5.date <- vect(era5.date, geom=c('lon', 'lat'), crs="+proj=longlat +datum=WGS84")
  empty.rast <- rast(nrows=15, ncol=41, extent=c(-85.125,-74.875,33.375,37.125), crs='+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0')
  era5.date <- rasterize(x=era5.date, y=empty.rast, field='t2m')
  era5.date <- flip(era5.date, direction="vertical")
  era5.date <- project(era5.date, crs(nc.borders))
  return(era5.date)
}

```

Map several date examples

```{r}
date <- as.Date('2022-06-06')
my.map.raster.points(era5.date(date, t2m.TN), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMIN-32)*5/9, 
                     date=date, title="TN North Carolina")
my.map.raster.points(era5.date(date, t2m.TX), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMAX-32)*5/9, 
                     date=date, title="TX North Carolina")

date <- as.Date('2022-06-14')
my.map.raster.points(era5.date(date, t2m.TN), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMIN-32)*5/9, 
                     date=date, title="TN North Carolina (very high TN)")
my.map.raster.points(era5.date(date, t2m.TX), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMAX-32)*5/9, 
                     date=date, title="TX North Carolina")

date <- as.Date('2022-06-20')
my.map.raster.points(era5.date(date, t2m.TN), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMIN-32)*5/9, 
                     date=date, title="TN North Carolina")
my.map.raster.points(era5.date(date, t2m.TX), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMAX-32)*5/9, 
                     date=date, title="TX North Carolina")

date <- as.Date('2022-07-26')
my.map.raster.points(era5.date(date, t2m.TN), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMIN-32)*5/9, 
                     date=date, title="TN North Carolina (very high TN) ")
my.map.raster.points(era5.date(date, t2m.TX), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMAX-32)*5/9, 
                     date=date, title="TX North Carolina")

date <- as.Date('2022-07-30')
my.map.raster.points(era5.date(date, t2m.TN), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMIN-32)*5/9, 
                     date=date, title="TN North Carolina (very high TN)")
my.map.raster.points(era5.date(date, t2m.TX), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMAX-32)*5/9, 
                     date=date, title="TX North Carolina")

date <- as.Date('2022-08-15')
my.map.raster.points(era5.date(date, t2m.TN), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMIN-32)*5/9, 
                     date=date, title="TN North Carolina (low TN)")
my.map.raster.points(era5.date(date, t2m.TX), 
                     points=aws[which(aws$DATE==date),], 
                     points.fill = (TMAX-32)*5/9, 
                     date=date, title="TX North Carolina")

```

```{r}
empty.rast <- rast(nrows=15, ncol=41, extent=c(-85.125,-74.875,33.375,37.125), 
                   crs='+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0')
date.vector <- as.vector(unique(sort(era5.t2m$date)))
era5.jja.vector.TN <- c()
era5.jja.vector.TX <- c()

for(d in date.vector){
  era5.d <- era5.t2m[which(era5.t2m$date==d),]
  era5.d.vect <- vect(era5.d, geom=c('lon', 'lat'), crs="+proj=longlat +datum=WGS84")
  
  era5.d.rast.TN <- rasterize(x=era5.d.vect, y=empty.rast, field=c('TN'))
  era5.d.rast.TN <- flip(era5.d.rast.TN, direction="vertical")
  era5.jja.vector.TN <- c(era5.jja.vector.TN, era5.d.rast.TN)
  
  era5.d.rast.TX <- rasterize(x=era5.d.vect, y=empty.rast, field=c('TX'))
  era5.d.rast.TX <- flip(era5.d.rast.TX, direction="vertical")
  era5.jja.vector.TX <- c(era5.jja.vector.TX, era5.d.rast.TX)
}

era5.jja.vector.TN <- rast(era5.jja.vector.TN)
names(era5.jja.vector.TN) <- date.vector
plot(subset(era5.jja.vector.TN, c(date.vector[1:5])))

era5.jja.vector.TX <- rast(era5.jja.vector.TX)
names(era5.jja.vector.TX) <- date.vector
plot(subset(era5.jja.vector.TX, c(date.vector[1:5])))

writeRaster(era5.jja.vector.TN, '../input/era5_daily_reanalysis_20220601_20220831_TN.tif')
writeRaster(era5.jja.vector.TX, '../input/era5_daily_reanalysis_20220601_20220831_TX.tif')
```

Session info

```{r}
sessionInfo()
```
