---
title: "space-time kriging"
output: html_document
date: "2023-08-29"
---

Libraries

```{r, message=F}
library(gstat)
library(data.table)
library(terra) 
library(sf)
library(sftime)
library(tidyverse)
library(ggplot2)
library(ggspatial)
library(tidyterra)
library(rgeos)
library(data.table) # -- for large flat datasets
library(DT)

# -- for timeseries
library(lubridate)
library(xts)
```

#### Open data

```{r}
obs <- fread("../input/NC-monitors-dailysummary-20220601-20220831-space-time-covariates.csv")
nc.borders <- vect("../input/NC_county_boundary/North_Carolina_State_and_County_Boundary_Polygons.shp")
```

Specific period

```{r}
p <- seq(as.Date('2022-07-20'), as.Date('2022-08-01'), by='1 day')
p <- as.character(p)

list.pred <- list()
for(d in p){
  pred.d <- fread(paste0("../input/prediction-grid/prediction_grid_points_urb_rur_space_time_covariates_", d, ".csv"))
  list.pred <- append(list.pred, list(pred.d))
}
pred.p <- rbindlist(list.pred, fill=TRUE)
obs.p <- obs[date %in% p,]
```

Convert to sftime and choose a projection in meters

```{r}
obs.p$date <- as.Date(obs.p$date)
obs.p.sft <- st_as_sftime(obs.p, 
                  coords=c('lon', 'lat'),
                  remove=F,
                  crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0", 
                  time_column_name = 'date')
pred.p$date <- as.Date(pred.p$date)
pred.p.sft <- st_as_sftime(pred.p, 
                    coords=c('lon', 'lat'),
                    remove=F,
                    crs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0", 
                    time_column_name = 'date')
 
crs.meters <- "+proj=lcc +lat_0=33.75 +lon_0=-79 +lat_1=36.1666666666667 +lat_2=34.3333333333333 +x_0=609601.219202439 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"

nc.borders <- project(nc.borders, crs.meters)
obs.p.sft <- st_transform(obs.p.sft, crs.meters)
pred.p.sft <- st_transform(pred.p.sft, crs.meters)

# -- store c(lon,lat) columns in crs.meters (instead of wgs84) 
obs.p.sft <- obs.p.sft %>%
    mutate(lon = unlist(map(obs.p.sft$geometry,1)),
           lat = unlist(map(obs.p.sft$geometry,2)))

pred.p.sft <- pred.p.sft %>%
    mutate(lon = unlist(map(pred.p.sft$geometry,1)),
           lat = unlist(map(pred.p.sft$geometry,2)))
```

Before removing NA, reconvert to data.table to easily create STFDF

```{r}
obs.p <- as.data.table(obs.p.sft)
obs.p <- obs.p[, geometry:=NULL]
```

Remove NA in obs.p.sft

```{r}
obs.p.sft <- obs.p.sft[!is.na(obs.p.sft$tmin) & !is.na(obs.p.sft$tmax),]
```

#### Variogram analysis

Convert DT to STFDF for gstat compatibility

```{r}
library(spacetime)
obs.p$date <- as.Date(obs.p$date)
obs.STFDF <- stConstruct(x=obs.p, space=c('lon', 'lat'), time='date', crs=CRS(crs.meters))
obs.STFDF <- as(obs.STFDF,'STFDF')
```

Directions: 0=north, 90=east, 180=south, 270=west

```{r}
cutoff=30000 
width=3000

plot(variogram(tmin ~ 1, obs.STFDF, cutoff=cutoff, width=width, tlags=0:8, alpha=180), wireframe=T, main='sample semivariogram')
```

TOBECONTINUEDDDDD

Variogram model

```{r}
cutoff <- 100000
v.emp <- variogram(tmin ~ 1, obs.p, cutoff=cutoff, width=width)

nugget <- 0.5
range <- 20000
vgm.sph <- vgm(psill=3, "Sph", range=range, nugget=0.5, cutoff=cutoff)
plot(v.emp, vgm.sph, cutoff=cutoff)

vgm.exp <- vgm(psill=3, "Exp", range=range, nugget=0.5, cutoff=cutoff)
plot(v.emp, vgm.exp, cutoff=cutoff)

vgm.mat <- vgm(psill=3, "Mat", range=range, nugget=0.5, cutoff=cutoff)
plot(v.emp, vgm.mat, cutoff=cutoff)
```

Fit Matern model

```{r}
v.fit <- fit.variogram(v.emp, vgm.mat, fit.range=F)
plot(v.emp, v.fit)
```

## Spatial kriging

#### Ordinary kriging

```{r}
cutoff <- 100000
nugget <- 0.5
range <- 20000
width <- 1000
v.emp <- variogram(tmin ~ 1, obs.p, cutoff=cutoff, width=width)
vgm.mat <- vgm(psill=3, "Mat", range=range, nugget=nugget, cutoff=cutoff)
v.fit <- fit.variogram(v.emp, vgm.mat, fit.range=F)
ok <- gstat::krige(tmin ~ 1, obs.p, newdata=pred.p, v.fit)
```

```{r}
m <- ggplot() +
  geom_spatvector(data = vect(ok), aes(color=var1.pred), size=.1) +
  geom_sf(data = obs.p, aes(geometry=geometry, color=tmin)) +
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "grey", linewidth=.3, fill = NA) +
  scale_color_whitebox_c(
      palette = "muted",
      labels = scales::label_number(suffix = "°C"),
      n.breaks = 12,
      guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Ordinary kriging"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
m
```

#### Universal kriging

```{r}
cutoff <- 100000
nugget <- 0.5
range <- 20000
width <- 1000
v.emp <- variogram(tmin ~ dem + imp + TN12am + tcc + build.fp, obs.p, cutoff=cutoff, width=width)
vgm.mat <- vgm(psill=2, "Mat", range=range, nugget=nugget, cutoff=cutoff)
plot(v.emp, vgm.mat)
```

```{r}
v.fit <- fit.variogram(v.emp, vgm.mat, fit.range=F)
plot(v.emp, v.fit)
uk <- gstat::krige(tmin ~ dem + imp + TN12am + tcc + build.fp, obs.p, newdata=pred.p, v.fit)
```

```{r}
m <- ggplot() +
  geom_spatvector(data = vect(uk), aes(color=var1.pred), size=.1) +
  geom_sf(data = obs.p, aes(geometry=geometry, color=tmin)) +
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "grey", linewidth=.3, fill = NA) +
  scale_color_whitebox_c(
      palette = "muted",
      labels = scales::label_number(suffix = "°C"),
      n.breaks = 12,
      guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Universal kriging"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
m
```

Same without ERA5

```{r}
cutoff <- 100000
nugget <- 0.5
range <- 20000
width <- 1000
v.emp <- variogram(tmin ~ dem + imp + tcc + build.fp, obs.p, cutoff=cutoff, width=width)
vgm.mat <- vgm(psill=3, "Mat", range=range, nugget=nugget, cutoff=cutoff)
plot(v.emp, vgm.mat)
v.fit <- fit.variogram(v.emp, vgm.mat, fit.range=F, fit.sill=F)
plot(v.emp, v.fit)
uk <- gstat::krige(tmin ~ dem + imp + tcc + build.fp, obs.p, newdata=pred.p, v.fit)
```

```{r}
plot.uk.noera5 <- ggplot() +
  geom_spatvector(data = vect(uk), aes(color=var1.pred), size=.1) +
  geom_sf(data = obs.p, aes(geometry=geometry, color=tmin)) +
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "grey", linewidth=.3, fill = NA) +
  scale_color_whitebox_c(
      palette = "muted",
      labels = scales::label_number(suffix = "°C"),
      n.breaks = 12,
      guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Universal kriging"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
plot.uk.noera5
```

Zoom on RTP area

```{r}
lat <- c(35.6, 36.11, 36.11, 35.6)
lon <- c(-79.19, -79.10, -78.39, -78.39)
ext.rtp <- vect(cbind(lon, lat), type="points", crs="+proj=longlat +datum=WGS84")
ext.rtp <- project(ext.rtp, crs.meters)
ext(ext.rtp)
```

```{r}
uk.rtp <- crop(vect(uk), ext.rtp)
imp.rtp <- crop(vect(pred.p[,'imp']), ext.rtp)
obs.rtp <- crop(vect(obs.p), ext.rtp)

ggplot() +
  geom_spatvector(data = uk.rtp, aes(color=var1.pred), size=.1) +
  geom_spatvector(data = obs.rtp, aes(color=tmin), shape=1, size=2) +
  #geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
  #             colour = "grey", linewidth=.3, fill = NA) +
  scale_color_whitebox_c(
      palette = "muted",
      labels = scales::label_number(suffix = "°C"),
      n.breaks = 12,
      guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Universal kriging RTP"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)

ggplot() +
  geom_spatvector(data = imp.rtp, aes(color=imp), size=.1) +
  geom_spatvector(data = obs.rtp, color='black', fill='black', shape=16, size=2) +
  scale_color_whitebox_c(
      palette = "muted",
      labels = scales::label_number(suffix = "%"),
      n.breaks = 12,
      guide = guide_legend(reverse = TRUE)
  ) +
  labs(
    fill = "",
    title = "Imperviousness RTP"
  ) +
  annotation_scale(
    location="bl", pad_x = unit(1, "cm"), 
		pad_y = unit(1, "cm"), 
		height = unit(0.30, "cm"), 
		text_cex = 1
	) +
	annotation_north_arrow(
	  location = "br", 
	  which_north = "true", 
		pad_x = unit(0.2, "cm"), 
		pad_y = unit(0.2, "cm")
	) +
	theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)
```

10-fold cross validation

```{r}
uk.loocv <- krige.cv(tmin ~ dem + imp + tcc + build.fp, obs.p, v.fit, nfold=nrow(obs.p))
uk.loocv
uk.10cv <- krige.cv(tmin ~ dem + imp + tcc + build.fp, obs.p, v.fit, nfold=10)
uk.10cv
```

RMSE computation

```{r}
RMSE.loocv <- sqrt(sum(uk.loocv$residual**2)/nrow(uk.loocv))
RMSE.10cv <- sqrt(sum(uk.10cv$residual**2)/nrow(uk.10cv))
cat('RMSE with loocv:', RMSE.loocv, '°C\n')
cat('RMSE with 10cv: ', RMSE.10cv, '°C')
```

Residuals

```{r}
ggplot(uk.loocv)+
  geom_point(aes(y=var1.pred, x=observed))+
  geom_abline(aes(slope=1, intercept=0), color='red')

ggplot(uk.10cv)+
  geom_point(aes(y=var1.pred, x=observed))+
  geom_abline(aes(slope=1, intercept=0), color='red')

pal.res <- RColorBrewer::brewer.pal(10, "RdBu")

ggplot()+
  geom_sf(data = uk.loocv, aes(geometry=geometry, color=residual)) +
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "grey", linewidth=.3, fill = NA) +
  scale_color_steps2(low='blue', mid='white', high='red', limits=c(-5,5), n.breaks=10)+
  theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)

ggplot()+
  geom_sf(data = uk.10cv, aes(geometry=geometry, color=residual)) +
  geom_sf(data = st_as_sf(nc.borders), aes(geometry=geometry), 
               colour = "grey", linewidth=.3, fill = NA) +
  scale_color_steps2(low='blue', mid='white', high='red', limits=c(-5,5), n.breaks=10)+
  theme(
			axis.text = element_text(size=12, family="serif"),
			plot.caption = element_text(size=10, family="serif"),
			legend.text = element_text(size=12, family="serif"),
			legend.title = element_text(size=12, family="serif"),
			panel.background = element_rect(fill = "white"),
			panel.grid.major=element_line(colour="grey")
		)

```

```{r}
head(uk.loocv)
ggplot()+
  geom_sf(data = uk.10cv, aes(geometry=geometry, color=fold))+
  
```
